---
title: "CCRVB-51: Franchini ALFQAvsALUM For MS"
author: "Kate Goldfarbmuren"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: FALSE
editor_options:
  
  chunk_output_type: console
params:  
  ############# where is data and analysis going
  ## these dirs should exist, usually on biowulf, but could be local
  ### parent folder where analysis output going, subfolders will be generated
  parent_dir: "~/../../Volumes/NCI_VB/franchini/ccrvb51_ALFQAvsALUM_KCG/ForMS" 
  ### where is input data mastertable
  input_dir: "~/../../Volumes/NCI_VB/franchini/ccrvb51_ALFQAvsALUM_KCG/ForMS/input_data" 

  ### overall study name, will be appended to every output
  study_name: "ccrvb51_ALFQAvsAlum"


  ### where is the original code file, don't change
  code_path: "~/../../Volumes/NCI_VB/goldfarbmurenkc/github/VaccineBranch/Franchini/ccrvb51_ALFQAvsALVAC_KCG.Rmd"
  #### where R libs on biowulf, shouldn't change
  lib_loc_to_use: "/data/goldfarbmurenkc/R/rhel8/4.3/"
  ### general functions, shouldn't change
  functions_path: "~/../../Volumes/NCI_VB/goldfarbmurenkc/github/VaccineBranch/general"
---


```{r setup, include=FALSE}
# more substantial setup for your document
knitr::opts_chunk$set(echo = FALSE, 
                      fig.width = 9, 
                      fig.height = 9, 
                      dev = "png", #change to cairo_pdf for pdf outputs
                      cache = FALSE,
                      warning = FALSE,
                      message = FALSE)


#https://github.com/rstudio/DT/issues/867
DT::datatable(
  matrix(), extensions="Buttons",
  options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))


```

```{r paths_lists, warning=FALSE, include=FALSE}
############################################################
# generate output directories
############################################################
parent_dir <- params$parent_dir
input_dir <- params$input_dir

#### generate
outputIMAGES_dir = paste0(parent_dir,"/IMAGES")
outputTABLES_dir = paste0(parent_dir,"/TABLES")
outputRData_dir = paste0(parent_dir,"/RData")
outputEnrichr_dir = paste0(parent_dir,"/TABLES/Enrichr")
output_dirs <- c(outputIMAGES_dir, outputTABLES_dir, outputRData_dir, outputEnrichr_dir)

for (curr_dir in output_dirs){
  if(!(file.exists(file.path(curr_dir)))){
    dir.create(file.path(curr_dir), showWarnings = TRUE)
  }
}
code_path = params$code_path
functions_path = params$functions_path
```


```{r source_pkgs, warning=FALSE, include=FALSE}
############################################################
# packages & functions
############################################################
# package list
list.of.packages=c("openxlsx","readxl", "tidyverse","cowplot", "pheatmap","ggpubr","survival","eulerr", "enrichR","ggsurvfit","OlinkAnalyze","knitr","broom","kableExtra","gtsummary",#"org.Hs.eg.db",
                   "gridExtra","DT","ggbeeswarm","ggrepel","plotly","ggplotify","flextable","PCAtools","officer","janitor",
                   #"DESeq2",
                   "RColorBrewer","colorspace","filesstrings","geepack","EnhancedVolcano",
                   "lmerTest",
                   "limma","edgeR","variancePartition","BiocParallel","statmod","corrplot","ggbiplot",
                   "strex","dendsort","dendextend","ggalluvial","rstatix",
                   "igraph")


#install as needed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

# source common functions, need to have NCI_VB mounted to run this
#source(file.path(functions_path,"CommonFunctions_KCG.R"))

# Rmd functions
#source(file.path(functions_path,'Rmd_functions_KCG.R'))

# load packages
invisible(lapply(list.of.packages, library, character.only = TRUE))
```




```{r loadData, warning=FALSE, include=FALSE}
############################################################
# master table file(s) to load
############################################################

######### read in file
xlsx_file_path <- file.path(params$input_dir,paste0("list_of_RAWdatasets.xlsx"))
tab_names<-openxlsx::getSheetNames(xlsx_file_path)
### list with separate element for each tab
masterlist <- lapply(tab_names,
                           openxlsx::read.xlsx,xlsxFile=xlsx_file_path)
names(masterlist) <- tab_names

phen_olink <- masterlist[["olink"]] %>%
  ### add in TOA for comparison
  dplyr::left_join(masterlist[["Challenge"]][,c("Sample_ID","TOA")], by = "Sample_ID") %>%
  dplyr::relocate(TOA, .after = "Time_Point")

olink_assay_columns <- setdiff(colnames(phen_olink),
                               c("animal_time_study","Study_number",
                               "Sample_ID","Time_Point","TOA","Protective_Vaccine"))


data_long <- masterlist[["otherAssays"]] %>%
  ### add in TOA for comparison
  dplyr::left_join(masterlist[["Challenge"]][,c("Sample_ID","TOA")], by = "Sample_ID") %>%
  dplyr::relocate(TOA, .after = "Time_Point") %>%

  ### make true long
  tidyr::pivot_longer(cols = !c(Sample_ID,Group,TOA,Time_Point),
                      names_to = "variable",
                      values_to = "value")
  

  #### prep list for stats & plotted values
list_of_stats <- list()
list_of_plotted <- list()

```


# {.tabset}

## Figure 1 (Infection) {.tabset}

### Fig 1bcd {.tabset}

Survival Curves

```{r Fig1b_setup}
#### set groups and order
fig_name_to_use <- "Fig1b"
group_var_to_use <- "Group"
groups_to_use <- c("Control","Alum")
colors_to_use <- c("black","blue")

### if need to trim out any groups
phen_long_to_use1 <- masterlist[["Challenge"]]


```

```{r TOA, results='asis', fig.height=6, fig.width=4}


        cat("#### ",paste0(fig_name_to_use, " "),"\n")
      cat('\n')
 
####### select animals for the aquistion curve
phen_long_to_use <- phen_long_to_use1 %>%
  dplyr::filter(!!rlang::sym(group_var_to_use) %in% groups_to_use) %>%
  dplyr::arrange(!!rlang::sym(group_var_to_use),Sample_ID) 


summ <- phen_long_to_use %>% 
  dplyr::mutate(status = recode(Infected, "0"=0,"1"=1), ## status needs to be 0,1
                time = as.numeric(TOA),
                vaccine = factor(!!rlang::sym(group_var_to_use), groups_to_use))


## survival function
#Surv(summ$time, summ$status) #calculates censored samples (i.e. that never had infection in time period observed)

### set title for 2 or 3 groups
if(length(groups_to_use)>2){
  title_to_use <- paste0(groups_to_use[1]," n=",dim(filter(summ, vaccine == groups_to_use[1]))[1],"; ",
                      groups_to_use[2]," n=",dim(filter(summ, vaccine == groups_to_use[2]))[1],"; ",
                      groups_to_use[3]," n=",dim(filter(summ, vaccine == groups_to_use[3]))[1])
}else{
   title_to_use <- paste0(groups_to_use[1]," n=",dim(filter(summ, vaccine == groups_to_use[1]))[1],"; ",
                      groups_to_use[2]," n=",dim(filter(summ, vaccine == groups_to_use[2]))[1])
 
}


sfit <- survfit2(Surv(time, status) ~ vaccine, data=summ) %>%
  ggsurvfit(linewidth=1)+
  scale_x_continuous(breaks=0:12, expand=c(0.02,0))+
  scale_y_continuous(label = scales::percent, breaks=seq(0,1, by=0.2), expand=c(0.015,0))+
  labs(title = title_to_use,
       x = "Weeks",
       y = "% uninfected macaques")+
  scale_color_manual(values=colors_to_use)+
  scale_fill_manual(values=colors_to_use)+
  add_confidence_interval()+
  add_pvalue(location = "annotation")+
  add_risktable(times=c(seq(0,12,1)))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")
        )


print(sfit)


 cat("\n")
  cat(' \n \n') ### this is the key!
  
  #### save data to list of plotted
  list_of_plotted[[paste0(fig_name_to_use)]] <- phen_long_to_use
  list_of_stats[[paste0(fig_name_to_use)]] <- sfit$data


```



```{r Fig1c_setup}
#### set groups and order
fig_name_to_use <- "Fig1c"
group_var_to_use <- "Group"
groups_to_use <- c("Control","ALFQA")
colors_to_use <- c("black","red")

### if need to trim out any groups
phen_long_to_use1 <- masterlist[["Challenge"]]

```

```{r TOA___Fig1c, ref.label=c('TOA'), results='asis', fig.height=6, fig.width=4}
```



```{r Fig1d_setup}
#### set groups and order
fig_name_to_use <- "Fig1d"
group_var_to_use <- "Group"
groups_to_use <- c("Alum","ALFQA")
colors_to_use <- c("blue","red")

### if need to trim out any groups
phen_long_to_use1 <- masterlist[["Challenge"]]

```

```{r TOA___Fig1d, ref.label=c('TOA'), results='asis', fig.height=6, fig.width=4}
```


### Fig 1e

```{r Fig1e, results='asis', fig.height=5, fig.width=7}

fig_name_to_use <- "Fig1e"
#### plot VL vs time by group, with ave, stdev & stats between groups for each timepoint





df_to_plot <- masterlist[["VL"]] %>%
  ### make longer
  tidyr::pivot_longer(!c(Group,Sample_ID), names_to = "Time_Point", values_to = "VL") %>%
  ## clean up TP
  dplyr::mutate(weeks_post_infection = as.numeric(str_after_last(Time_Point,"\\.")),
                log10_VL = log10(VL),
                ## factorize group
                Group = factor(Group, c("Control","Alum","ALFQA"))) 

### get geo mean & 95% CI
df_summary <- df_to_plot %>%
  group_by(Group,weeks_post_infection) %>%
  dplyr::summarise(
    sd = sd(log10_VL, na.rm = TRUE),
    simp_mean = mean(log10_VL, na.rm = TRUE),
    geo_mean = exp(mean(log(log10_VL[log10_VL>0]), na.rm=TRUE)),
    N = n(),
    
  ) %>%
  dplyr::mutate(se = sd/sqrt(N),
                ciMult = qt(0.95/2 + .5, N-1), ### 95% confidence interval multiplier
                ci = se*ciMult)
  
#### get MW pvalues
stat_test <- df_to_plot %>%
  group_by(weeks_post_infection) %>%
  wilcox_test(log10_VL ~ Group) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p") %>%
  ### remove ALFQA vs Alum
  dplyr::filter(group1 == "Control" & group2 == "Alum" |
                  group1 == "Control" & group2 == "ALFQA" ) %>%
  ## sort 
 # dplyr::arrange(group2) %>%
  add_xy_position(x = "weeks_post_infection", dodge = 0.3) %>%
  ### fix x position
  dplyr::mutate(x = weeks_post_infection,
                xmin = weeks_post_infection,
                xmax = weeks_post_infection)


p <- ggplot(df_summary,
            aes(x = weeks_post_infection, y = geo_mean))+
  geom_errorbar(
    aes(ymin = geo_mean-ci, ymax = geo_mean+ci, color = Group),
    position = position_dodge(0.3), width = 0.2
    )+
  geom_line(aes(group = Group, color = Group), position = position_dodge(0.3))+
  geom_point(aes(color = Group), position = position_dodge(0.3)) +
  scale_color_manual(values = c(Control = "black",
                                Alum = "blue",
                                ALFQA = "red"))+
  scale_x_continuous(breaks = seq(0,12,2), limits = c(0,13), expand = c(0,0))+
  scale_y_continuous(breaks = seq(1,10,1), limits = c(1,10), expand = c(0,0))+
  ylab("log10(VL)")+
  stat_pvalue_manual(stat_test,  
                     label = "p.signif", tip.length = 0,
                     remove.bracket = TRUE,
                     color = "group2"
  )+
  theme_bw()+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) +
  ## remove a from legend
  guides(
  color = guide_legend(
    override.aes = aes(label = "")
  ))

plot(p)


cat("\n")
  cat(' \n \n') ### this is the key!
  
 

    #### save data to list of plotted
  list_of_plotted[[paste0(fig_name_to_use)]] <- df_to_plot %>%
    ### add summary stats
    dplyr::left_join(df_summary, by = c("Group","weeks_post_infection"))
  
  list_of_stats[[paste0(fig_name_to_use)]] <- stat_test

 

```


## Supp Figure 1 (Viral load) {.tabset}

Scatterplots vs 3 groups

```{r SuppFig1_scatterplots, results='asis', fig.height=4, fig.width=4}

fig_name_to_use <- "SuppFig1"
#### plot VL vs time by group, with ave, stdev & stats between groups for each timepoint
dodge_to_use <- 0.8

df_to_plot_prep <- masterlist[["VL"]] %>%
  ### make longer
  tidyr::pivot_longer(!c(Group,Sample_ID), names_to = "Time_Point", values_to = "VL") %>%
  ## clean up TP
  dplyr::mutate(weeks_post_infection = as.numeric(str_after_last(Time_Point,"\\.")),
                log10_VL = log10(VL),
                ## factorize group
                Group = factor(Group, c("Control","Alum","ALFQA"))) 

### wide format for plotting
df_to_plot <- df_to_plot_prep %>%
  ## remove non log10
  dplyr::select(-VL,-Time_Point) %>%
  ## make wider
  dplyr::mutate(weeks_post_infection = paste0("VL_week",weeks_post_infection)) %>%
  tidyr::pivot_wider(names_from = "weeks_post_infection",
                     values_from = "log10_VL")

### get geo mean (same as normal mean) & 95% CI
df_summary <- df_to_plot_prep %>%
  group_by(Group,weeks_post_infection) %>%
  dplyr::summarise(
    sd = sd(log10_VL, na.rm = TRUE),
    mean = mean(log10_VL, na.rm = TRUE),
    geo_mean = exp(mean(log(log10_VL[log10_VL>0]), na.rm=TRUE)),
    N = n(),
    
  ) %>%
  dplyr::mutate(se = sd/sqrt(N),
                ciMult = qt(0.95/2 + .5, N-1), ### 95% confidence interval multiplier
                ci = se*ciMult)
  
#### get MW pvalues
stat_test <- df_to_plot_prep %>%
  group_by(weeks_post_infection) %>%
  wilcox_test(log10_VL ~ Group) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p") 

### set up df to collect stats
stat_test <- data.frame()

#### list of plots with each variable separately
list_of_plots <- list()
for (i in unique(df_to_plot_prep$weeks_post_infection)){
 
  #### get stats formatting
#  https://www.datanovia.com/en/blog/how-to-add-p-values-onto-a-grouped-ggplot-using-the-ggpubr-r-package/
 
  
        cat("### ",paste0("week ",i, " post infection "),"\n")
      cat('\n')
 
  ###### p values for this week
   curr_stat_test <- df_to_plot_prep %>%
     ## filter to this week
     dplyr::filter(weeks_post_infection == i) %>%
  group_by(weeks_post_infection) %>%
  wilcox_test(log10_VL ~ Group) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p")  %>%
     dplyr::filter(weeks_post_infection == i) %>%

  add_xy_position(x = "weeks_post_infection", dodge = dodge_to_use) %>%
     ### then replace x positions since aren't useful
     dplyr::mutate(x = 2, 
                   xmin = ifelse(group1 == "Control",1,2),
                   xmax = ifelse(group2 == "ALFQA",3,2))
     
     
  
  dodge_to_use <- 0.8
  #### base plot (needs to NOT be vanilla ggplot)
  p1 <- ggbarplot(df_to_plot, x = "Group", y = paste0("VL_week",i), color = "Group", fill = "Group",
                  add = "mean", position = position_dodge(dodge_to_use), alpha = 0.3) + 
  scale_color_manual(values = c(Control = "black",
                                Alum = "blue",
                                ALFQA = "red"))+
  scale_fill_manual(values = c(Control = "black",
                                Alum = "blue",
                                ALFQA = "red"))+
    geom_beeswarm(aes(x = Group, y = !!rlang::sym(paste0("VL_week",i)), group = Group, fill = Group), dodge.width = dodge_to_use, cex = 3, size = 12, shape = 21, color = "black") + 
    
    # geom_bar(aes(color = Group_f),stat = "summary", fun = "mean", na.rm = TRUE,alpha = 0.3,
    #    position = position_dodge(width = dodge_to_use), width = 0.7) +
    #  geom_errorbar(stat = "summary", fun.data = "mean_sdl", 
    #         fun.args = list(mult = 1),
    #         position =  position_dodge(width = dodge_to_use), width = 0.5) +
    # 
  
    #### add pvalues between gropus
    stat_pvalue_manual(
  curr_stat_test,  label = "p", tip.length = 0
  )+
    stat_summary(
    fun = "mean",        #argument updated in new version.
    geom = "point",
    col = "grey30",
    size = 3,
    shape = 95,
    fill = "grey"
  )+

   
    scale_y_continuous(breaks = seq(1,10,1), limits = c(1,10),
                       expand = expansion(mult = c(0, 0.1))
                       )+
    ylab(paste0("log10(VL week",i,")"))+

      theme_bw()+
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.title.x = element_blank(),
            axis.title.y = element_text(size = 9),
            axis.line = element_line(colour = "black"),
            panel.border = element_blank(),
    panel.background = element_blank())
    
plot(p1)
    
    
 ### add to stat_test
   stat_test <- rbind(stat_test, curr_stat_test)
  
  
  cat("\n")
  cat(' \n \n') ### this is the key!
  
}


  

 cat("\n")
  cat(' \n \n') ### this is the key!
  
  
    #### save data to list of plotted
  list_of_plotted[[paste0(fig_name_to_use)]] <- df_to_plot 
  
  list_of_stats[[paste0(fig_name_to_use)]] <- stat_test

 

```


## Figure 2 (Functional Assays in Blood) {.tabset}


```{r scatterVsGroupSetup___Fig2acefghjk}
composite_fig_name <- "Fig2acefghjk"
scatter_df <- data.frame("name"=c("Fig2a",
                                  "Fig2c",
                                  "Fig2e",
                                  "Fig2f",
                                  "Fig2g",
                                  "Fig2h",
                                  "Fig2j",
                                  "Fig2k"),
                         "y"=c("Plasma_peptide26gp120_antibody_response__WkPost",
                               "Vaginal_secretions_dV1gp120_antibody_titers__WkPost",
                               "Plasma_Trogocytosis_dV1gp120__Change",
                               "Plasma_ADCP_SIVgp120dV1__Change",
                               "Plasma_ADNP_SIVgp120dV1__Change",
                               "Plasma_ADNP_SIVgp120WT__Change",
                               "PBMCs_PctCD14_efferocytes__Change",
                               "PBMCs_engulfing_capability__Change"),
                         "trans_y"=c("identity",
                                     "log10",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other"))


fill_colors_to_use <- c(ALFQA="red",Alum="blue")
outline_colors_to_use <- c("black","black")

```

```{r scatterplotsVsGroupPlot, results='asis', fig.height=3, fig.width=4}


 #### if only one pair
  if(dim(scatter_df)[1] == 1){
    
             cat(" ",paste0(composite_fig_name),"\n")
      cat('\n')
 
    
  }else{
    #### more than one, then start with group name in nested tabset
    
           cat("### ",paste0(composite_fig_name, " {.tabset}"),"\n")
      cat('\n')
 
  }

#### for each pair to plot
for(curr_pair in 1:nrow(scatter_df)){
  
  
  #### if only one pair
  if(dim(scatter_df)[1] == 1){
    
           cat("### ",paste0(scatter_df$name[curr_pair]),"\n")
      cat('\n')
 
  }else{
    #### more than one, then in nested tabset
    
           cat("#### ",paste0(scatter_df$name[curr_pair]),"\n")
      cat('\n')
 
  }
  

  
      if(scatter_df$dataset[curr_pair] == "other"){
        curr_phen <- data_long %>%
          ## combine variable/timepoint
          dplyr::mutate(tp_simp = ifelse(grepl("Wk",Time_Point),
                                         "WkPost",Time_Point),
                        var_time = paste0(variable,"__",tp_simp),
                        variable = NULL,
                        Time_Point = NULL,
                        tp_simp = NULL,
                        TOA = NULL ### not needed for this plot
                        ) %>%
          ### collapse duplicates with NA
          dplyr::group_by(Sample_ID,Group,var_time) %>%
          dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value") 
      }else{
        if(scatter_df$dataset[curr_pair] == "olink"){
          curr_phen <- curr_phen_olink %>%
            ## trim to only columns that need
            dplyr::select(c(group,TOA,Sample_ID,Time_Point,!!!rlang::syms(olink_assay_columns_new))) %>%
            ## make longer
            tidyr::pivot_longer(cols = !c(group,TOA,Sample_ID,Time_Point), names_to = "variable", values_to = "value") %>%
                      ## combine variable/timepoint
          dplyr::mutate(var_time = paste0(variable,"__",Time_Point),
                        variable = NULL,
                        Time_Point = NULL,
                        Group = ifelse(group == "ALUM","Alum",group),
                        TOA = NULL ### not needed for this plot
                        ) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value")

        }else{
          if(scatter_df$dataset[curr_pair] == "otherStudy"){
            
            curr_phen <- masterlist[["otherStudy"]] %>%
              dplyr::mutate(Group = factor(Group, c("Alum","ALFQA")))
            
          }else{
          
          #### for both (not used for vs TOA)
          
          ## first find animals that are in both datasets
overlap_animals <- intersect(unique(data_long$Sample_ID),unique(phen_olink$Sample_ID)) ### 41 animals

          curr_phen <- data_long %>%
          ## combine variable/timepoint
          dplyr::mutate(tp_simp = ifelse(grepl("Wk",Time_Point),
                                         "WkPost",Time_Point),
                        var_time = paste0(variable,"__",tp_simp),
                        variable = NULL,
                        Time_Point = NULL,
                        tp_simp = NULL,
                        TOA = NULL ### not needed for this plot
                        ) %>%
          ### collapse duplicates with NA
          dplyr::group_by(Sample_ID,Group,var_time) %>%
          dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value") %>%
  ### trim to overlapping animals
  dplyr::filter(Sample_ID %in% overlap_animals) %>%
  ####### add in olink
  dplyr::full_join(phen_olink %>%
                     ### trim to overlapping animals
                     dplyr::filter(Sample_ID %in% overlap_animals) %>%
                     ### combine variable & timepoint for olink assays
                     ## first need to make longer
                     tidyr::pivot_longer(cols = c(!!!rlang::syms(olink_assay_columns_new)),
                                         names_to = "variable",
                                         values_to = "value") %>%
                     dplyr::mutate(var_time = paste0(variable,"__",Time_Point)) %>%
                     ### remove anything with timepoint
                     dplyr::select(Sample_ID,var_time,value) %>%
                     tidyr::pivot_wider(names_from = "var_time",values_from = "value")
                   ,
                   by = c("Sample_ID"))
          }
        }
      }
      

    df_for_plotting <- curr_phen %>%
          ### trim to only these vars
          dplyr::select(!!rlang::sym(scatter_df$y[curr_pair]),Group,Sample_ID) %>%
      ### filter to only animals with data for this assay
      dplyr::filter(!is.na(!!rlang::sym(scatter_df$y[curr_pair]))) %>%
      ### factorize Group
      dplyr::mutate(Group = factor(Group, c("Alum","ALFQA")))


      if(scatter_df$trans_y[curr_pair] == "identity"){
        ylab_to_use <- paste0(str_before_first(scatter_df$y[curr_pair],"__"),"\n",str_after_first(scatter_df$y[curr_pair],"__"))
      }else{
        ylab_to_use <- paste0(scatter_df$trans_y[curr_pair],"(",str_before_first(scatter_df$y[curr_pair],"__"),")","\n",str_after_first(scatter_df$y[curr_pair],"__") )
      }
      

        #### plot scatterplot with both groups
        p <- ggplot(df_for_plotting,
                    aes(x = Group, y = !!rlang::sym(scatter_df$y[curr_pair]), fill = Group, color = Group))+
          geom_quasirandom(width = 0.4, shape = 21, size = 3)+
          scale_fill_manual(values = fill_colors_to_use)+
          scale_color_manual(values = outline_colors_to_use)+          
          #### add MW stats
          stat_compare_means(method = "wilcox.test", label.x = 1.25, label="p.format")+
          ### add mean
          stat_summary(color = "black", geom = "point", shape = 95, size = 12, show.legend = F,
                     fun = mean)+
          ### add horizontal dotted line at zero
          geom_hline(yintercept = 0, linetype = "dashed", color = "grey")+
          

          ylab(ylab_to_use)+
          scale_y_continuous(transform = scatter_df$trans_y[curr_pair])+  

          theme_bw()+
              theme(
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.title = element_text(size = 10),
            panel.border = element_blank(), axis.line = element_line(colour = "black")
        )+
            ## remove a from legend
  guides(
  color = guide_legend(
    override.aes = aes(label = "")
  ))
        
        plot(p)
        
          
        cat("\n")
    cat(' \n \n') ### this is the key!
    
    
    #### save
        list_of_plotted[[paste0(scatter_df$name[curr_pair])]] <- df_for_plotting %>%
          dplyr::relocate(Group,Sample_ID) %>%
          ##sort by group, sampleID
          dplyr::arrange(Group,Sample_ID) 
          
          

  
}
          
        cat("\n")
    cat(' \n \n') ### this is the key!

    
     
```




```{r scatterVsOtherSetup___Fig2b}
composite_fig_name <- ""
scatter_df <- data.frame("name"=c("Fig2b"),
                         "x"=c("Plasma_peptide26gp120_antibody_response__WkPost"),
                         "trans_x"=c("identity"),
                         "y"=c("Plasma_AvidityScore_cV2gp120__WkPost"),
                         "trans_y"=c("log10"),
                         "dataset"=c("other"))


```

```{r scatterplotsVsOtherPlot, results='asis', fig.height=4, fig.width=5}


 #### if only one pair
  if(dim(scatter_df)[1] == 1){
    
             cat(" ",paste0(composite_fig_name),"\n")
      cat('\n')
 
    
  }else{
    #### more than one, then start with group name in nested tabset
    
           cat("### ",paste0(composite_fig_name, " {.tabset}"),"\n")
      cat('\n')
 
  }

#### for each pair to plot
for(curr_pair in 1:nrow(scatter_df)){
  
  
  #### if only one pair
  if(dim(scatter_df)[1] == 1){
    
           cat("### ",paste0(scatter_df$name[curr_pair]),"\n")
      cat('\n')
 
  }else{
    #### more than one, then in nested tabset
    
           cat("#### ",paste0(scatter_df$name[curr_pair]),"\n")
      cat('\n')
 
  }
  
    
           cat(" ",paste0(scatter_df$x[curr_pair]," vs ",scatter_df$y[curr_pair]),"\n")
      cat('\n')
  
      if(scatter_df$dataset[curr_pair] == "other"){
        curr_phen <- data_long %>%
          ## combine variable/timepoint
          dplyr::mutate(tp_simp = ifelse(grepl("Wk",Time_Point),
                                         "WkPost",Time_Point),
                        var_time = paste0(variable,"__",tp_simp),
                        variable = NULL,
                        Time_Point = NULL,
                        tp_simp = NULL) %>%
          ### collapse duplicates with NA
          dplyr::group_by(Sample_ID,Group,TOA,var_time) %>%
          dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value")
      }else{
        if(scatter_df$dataset[curr_pair] == "olink"){
          curr_phen <- curr_phen_olink %>%
            ## trim to only columns that need
            dplyr::select(c(group,TOA,Sample_ID,Time_Point,!!!rlang::syms(olink_assay_columns_new))) %>%
            ## make longer
            tidyr::pivot_longer(cols = !c(group,TOA,Sample_ID,Time_Point), names_to = "variable", values_to = "value") %>%
                      ## combine variable/timepoint
          dplyr::mutate(var_time = paste0(variable,"__",Time_Point),
                        variable = NULL,
                        Time_Point = NULL,
                        Group = group) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value")

        }else{
          #### for both
          
          ## first find animals that are in both datasets
overlap_animals <- intersect(unique(data_long$Sample_ID),unique(phen_olink$Sample_ID)) ### 41 animals

          curr_phen <- data_long %>%
          ## combine variable/timepoint
          dplyr::mutate(tp_simp = ifelse(grepl("Wk",Time_Point),
                                         "WkPost",Time_Point),
                        var_time = paste0(variable,"__",tp_simp),
                        variable = NULL,
                        Time_Point = NULL,
                        tp_simp = NULL) %>%
          ### collapse duplicates with NA
          dplyr::group_by(Sample_ID,Group,TOA,var_time) %>%
          dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value") %>%
  ### trim to overlapping animals
  dplyr::filter(Sample_ID %in% overlap_animals) %>%
  ####### add in olink
  dplyr::full_join(phen_olink %>%
                     ### trim to overlapping animals
                     dplyr::filter(Sample_ID %in% overlap_animals) %>%
                     ### combine variable & timepoint for olink assays
                     ## first need to make longer
                     tidyr::pivot_longer(cols = c(!!!rlang::syms(olink_assay_columns_new)),
                                         names_to = "variable",
                                         values_to = "value") %>%
                     dplyr::mutate(var_time = paste0(variable,"__",Time_Point)) %>%
                     ### remove anything with timepoint
                     dplyr::select(Sample_ID,var_time,value) %>%
                     tidyr::pivot_wider(names_from = "var_time",values_from = "value")
                   ,
                   by = c("Sample_ID"))
        }
      }
      

    df_for_plotting <- curr_phen %>%
          ### trim to only these vars
          dplyr::select(!!rlang::sym(scatter_df$x[curr_pair]),!!rlang::sym(scatter_df$y[curr_pair]),Group,Sample_ID) %>%
            ### filter to only animals with data for these assays
      dplyr::filter(!is.na(!!rlang::sym(scatter_df$y[curr_pair])) & !is.na(!!rlang::sym(scatter_df$x[curr_pair]))) 

      if(scatter_df$trans_x[curr_pair] == "identity"){
        xlab_to_use <- paste0(str_before_first(scatter_df$x[curr_pair],"__"),"\n",str_after_first(scatter_df$x[curr_pair],"__"))
      }else{
        xlab_to_use <- paste0(scatter_df$trans_x[curr_pair],"(",str_before_first(scatter_df$x[curr_pair],"__"),")","\n",str_after_first(scatter_df$x[curr_pair],"__") )
      }

           if(scatter_df$trans_y[curr_pair] == "identity"){
        ylab_to_use <- paste0(str_before_first(scatter_df$y[curr_pair],"__"),"\n",str_after_first(scatter_df$y[curr_pair],"__"))
      }else{
        ylab_to_use <- paste0(scatter_df$trans_y[curr_pair],"(",str_before_first(scatter_df$y[curr_pair],"__"),")","\n",str_after_first(scatter_df$y[curr_pair],"__") )
      }
      

        #### plot scatterplot with both groups
        p <- ggplot(df_for_plotting,
                    aes(x = !!rlang::sym(scatter_df$x[curr_pair]), y = !!rlang::sym(scatter_df$y[curr_pair]), fill = Group))+
          geom_point(shape = 21, size = 3)+
          scale_fill_manual(values = c("red","blue"))+
          ### add spearman stats
          stat_cor(aes(color = Group), method = "spearman", show.legend = FALSE)+
          ### add reg line
          geom_smooth(aes(x = as.numeric(!!rlang::sym(scatter_df$x[curr_pair])), color = Group), method = lm, se = FALSE, linewidth = 0.5, formula = 'y ~ x ') +
         scale_x_continuous(transform = scatter_df$trans_x[curr_pair])+  
          xlab(xlab_to_use)+
          ylab(ylab_to_use)+
          scale_y_continuous(transform = scatter_df$trans_y[curr_pair])+  
          scale_color_manual(values = c("red","blue"))+
          theme_bw()+
              theme(
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.title = element_text(size = 10),
            panel.border = element_blank(), axis.line = element_line(colour = "black")
        )
        
        plot(p)
        
          
        cat("\n")
    cat(' \n \n') ### this is the key!
    
    
    #### save
        list_of_plotted[[paste0(scatter_df$name[curr_pair])]] <- df_for_plotting %>%
          dplyr::relocate(Group,Sample_ID) %>%
          ##sort by group, sampleID
          dplyr::arrange(Group,Sample_ID)

  
}
          
        cat("\n")
    cat(' \n \n') ### this is the key!

    
     
```



```{r scatterVsTOASetup___Fig2di}
composite_fig_name <- "Fig2di"
scatter_df <- data.frame("name"=c("Fig2d",
                                  "Fig2i"),
                         "y"=c("Plasma_ADCC_Killing__WkPost",
                               "Plasma_ADNP_SIVgp120dV1__Change"),
                         "trans_y"=c("identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other"))

```

```{r scatterplotsVsTOAPlot, results='asis', fig.height=4, fig.width=5}


 #### if only one pair
  if(dim(scatter_df)[1] == 1){
    
             cat(" ",paste0(composite_fig_name),"\n")
      cat('\n')
 
    
  }else{
    #### more than one, then start with group name in nested tabset
    
           cat("### ",paste0(composite_fig_name, " {.tabset}"),"\n")
      cat('\n')
 
  }

#### for each pair to plot
for(curr_pair in 1:nrow(scatter_df)){
  
  
  #### if only one pair
  if(dim(scatter_df)[1] == 1){
    
           cat("### ",paste0(scatter_df$name[curr_pair]),"\n")
      cat('\n')
 
  }else{
    #### more than one, then in nested tabset
    
           cat("#### ",paste0(scatter_df$name[curr_pair]),"\n")
      cat('\n')
 
  }
  

  
      if(scatter_df$dataset[curr_pair] == "other"){
        curr_phen <- data_long %>%
          ## combine variable/timepoint
          dplyr::mutate(tp_simp = ifelse(grepl("Wk",Time_Point),
                                         "WkPost",Time_Point),
                        var_time = paste0(variable,"__",tp_simp),
                        variable = NULL,
                        Time_Point = NULL,
                        tp_simp = NULL) %>%
          ### collapse duplicates with NA
          dplyr::group_by(Sample_ID,Group,TOA,var_time) %>%
          dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value")
      }else{
        if(scatter_df$dataset[curr_pair] == "olink"){
          curr_phen <- curr_phen_olink %>%
            ## trim to only columns that need
            dplyr::select(c(group,TOA,Sample_ID,Time_Point,!!!rlang::syms(olink_assay_columns_new))) %>%
            ## make longer
            tidyr::pivot_longer(cols = !c(group,TOA,Sample_ID,Time_Point), names_to = "variable", values_to = "value") %>%
                      ## combine variable/timepoint
          dplyr::mutate(var_time = paste0(variable,"__",Time_Point),
                        variable = NULL,
                        Time_Point = NULL,
                        Group = group) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value")

        }else{
          #### for both (not used for vs TOA)
          
          ## first find animals that are in both datasets
overlap_animals <- intersect(unique(data_long$Sample_ID),unique(phen_olink$Sample_ID)) ### 41 animals

          curr_phen <- data_long %>%
          ## combine variable/timepoint
          dplyr::mutate(tp_simp = ifelse(grepl("Wk",Time_Point),
                                         "WkPost",Time_Point),
                        var_time = paste0(variable,"__",tp_simp),
                        variable = NULL,
                        Time_Point = NULL,
                        tp_simp = NULL) %>%
          ### collapse duplicates with NA
          dplyr::group_by(Sample_ID,Group,TOA,var_time) %>%
          dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
          ### make wider
          tidyr::pivot_wider(names_from = "var_time", values_from = "value") %>%
  ### trim to overlapping animals
  dplyr::filter(Sample_ID %in% overlap_animals) %>%
  ####### add in olink
  dplyr::full_join(phen_olink %>%
                     ### trim to overlapping animals
                     dplyr::filter(Sample_ID %in% overlap_animals) %>%
                     ### combine variable & timepoint for olink assays
                     ## first need to make longer
                     tidyr::pivot_longer(cols = c(!!!rlang::syms(olink_assay_columns_new)),
                                         names_to = "variable",
                                         values_to = "value") %>%
                     dplyr::mutate(var_time = paste0(variable,"__",Time_Point)) %>%
                     ### remove anything with timepoint
                     dplyr::select(Sample_ID,var_time,value) %>%
                     tidyr::pivot_wider(names_from = "var_time",values_from = "value")
                   ,
                   by = c("Sample_ID"))
        }
      }
      

    df_for_plotting <- curr_phen %>%
          ### trim to only these vars
          dplyr::select(TOA,!!rlang::sym(scatter_df$y[curr_pair]),Group,Sample_ID) %>%
            ### filter to only animals with data for these assays
      dplyr::filter(!is.na(!!rlang::sym(scatter_df$y[curr_pair])) & !is.na(TOA))


      if(scatter_df$trans_y[curr_pair] == "identity"){
        ylab_to_use <- paste0(str_before_first(scatter_df$y[curr_pair],"__"),"\n",str_after_first(scatter_df$y[curr_pair],"__"))
      }else{
        ylab_to_use <- paste0(scatter_df$trans_y[curr_pair],"(",str_before_first(scatter_df$y[curr_pair],"__"),")","\n",str_after_first(scatter_df$y[curr_pair],"__") )
      }
      

        #### plot scatterplot with both groups
        p <- ggplot(df_for_plotting,
                    aes(x = TOA, y = !!rlang::sym(scatter_df$y[curr_pair]), fill = Group))+
          geom_point(shape = 21, size = 3)+
          scale_fill_manual(values = c("red","blue"))+
          ### add spearman stats
          stat_cor(aes(color = Group), method = "spearman", show.legend = FALSE)+
          ### add reg line
          geom_smooth(aes(x = as.numeric(TOA), color = Group), method = lm, se = FALSE, linewidth = 0.5, formula = 'y ~ x ') +
         scale_x_continuous(breaks = seq(0,12,1), limits = c(0,12))+  
          xlab("N. vaginal challenges")+
          ylab(ylab_to_use)+
          scale_y_continuous(transform = scatter_df$trans_y[curr_pair])+  
          scale_color_manual(values = c("red","blue"))+
          theme_bw()+
              theme(
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.title = element_text(size = 10),
            panel.border = element_blank(), axis.line = element_line(colour = "black")
        )
        
        plot(p)
        
          
        cat("\n")
    cat(' \n \n') ### this is the key!
    
    
    #### save
        list_of_plotted[[paste0(scatter_df$name[curr_pair])]] <- df_for_plotting %>%
          dplyr::relocate(Group,Sample_ID) %>%
          ##sort by group, sampleID
          dplyr::arrange(Group,Sample_ID)

  
}
          
        cat("\n")
    cat(' \n \n') ### this is the key!

    
     
```





```{r setup___Fig2l}
fig_name_to_use <- "Fig2l"
assays_to_use <- sort(unique(data_long$variable)[!grepl("RectMucosa", unique(data_long$variable))])
```

### `r paste0(fig_name_to_use)` {.tabset}

```{r setSamples_other}
### curr_phen has rows = animals/time, columns = assays
curr_phen <- data_long %>%
  ## make simpler timepoint
  dplyr::mutate(Time_simp = ifelse(grepl("Wk",Time_Point),"WkPost",Time_Point),
                Time_Point = Time_simp,
                Time_simp = NULL,
                ### also make group with all caps
                group = toupper(Group)) %>%
  ## make wider
  unique() %>%
  ## remove duplicate rows with NA + number
  dplyr::group_by(Sample_ID,Group, group,TOA,Time_Point,variable) %>%
  dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
  ## TRIM TO ONLY vars of interest
  dplyr::filter(variable %in% assays_to_use) %>%
  tidyr::pivot_wider(names_from = "variable",values_from = "value") %>%
    ### remove rows with all missing 
  dplyr::filter_at(vars(assays_to_use), any_vars(!is.na(.))) 


### save values used to list_of_plotted
list_of_plotted[[paste0(fig_name_to_use)]] <- curr_phen %>%
  dplyr::arrange(group,Sample_ID) %>%
  ## remove duplicate Group
  dplyr::select(-Group)


```

```{r MW_2group_other}
#Mann-Whitney/Wilcoxon test between groups at each timepoint
#### get data into true wide format
## rows = animals (+ group and any metadata variables that are intrinsic (i.e. don't change with time) to the animal)
## columns = variables + time
group_levels <- c("Alum","ALFQA")

data_wide <- data_long %>%
  ### combine variable + time 
  dplyr::mutate(var_time = paste0(variable,"__",Time_Point),
                ## remove variable and time
                variable = NULL, 
                Time_Point = NULL) %>%
  ### make wider
  tidyr::pivot_wider(names_from = "var_time", values_from = "value") %>%
  ### factorize group
  dplyr::mutate(Group_f = factor(Group, group_levels),
                ## also make rownames be animal
                rowname = Sample_ID) %>%
  column_to_rownames()


###### trim to only variables with some variation (i.e. if all animals == 0 then removed)
data_trim <- data_wide %>%
  dplyr::select(where(~n_distinct(.) > 1))


        ### trim to only the variables with data for both groups (removes if all NA in one group, even if there is data in the other group because will have nothing to compare to)
        trim_df1 <- data_trim[which(data_trim$Group_f == group_levels[1]),] #%>% dplyr::select(-Group_f)
        columns_with_noNAs1 <- colnames(trim_df1)[which(colSums(is.na(trim_df1)) == 0)]
        columns_with_someNAs1 <- colnames(trim_df1)[which(colSums(is.na(trim_df1)) > 0 & colSums(is.na(trim_df1)) < dim(trim_df1)[1])]
        columns_with_allNAs1 <- colnames(trim_df1)[which(colSums(is.na(trim_df1)) == dim(trim_df1)[1])]
   
        
        trim_df2 <- data_trim[which(data_trim$Group_f == group_levels[2]),] #%>% dplyr::select(-Group_f)
        columns_with_noNAs2 <- colnames(trim_df2)[which(colSums(is.na(trim_df2)) == 0)]
        columns_with_someNAs2 <- colnames(trim_df2)[which(colSums(is.na(trim_df2)) > 0 & colSums(is.na(trim_df2)) < dim(trim_df2)[1])]
        columns_with_allNAs2 <- colnames(trim_df2)[which(colSums(is.na(trim_df2)) == dim(trim_df2)[1])]



        
      new_curr_df <- data_trim[,intersect(c(columns_with_noNAs1,columns_with_someNAs1),c(columns_with_noNAs2,columns_with_someNAs2))] 
      
      
        ## generate matrix with only variables that you want to test (no character variables and nothing that has zero variation across all animals)
      Assay_mat <- data.matrix(new_curr_df %>%
                               dplyr::select(-c(Sample_ID,Group,Group_f,TOA)))
      
      
          curr_p_raw <- apply(Assay_mat, 2, function(x) wilcox.test(x ~ Group_f, data = new_curr_df)$p.value)
          curr_p_adj <- p.adjust(curr_p_raw, method = 'BH') 
          # curr_est <- apply(new_Assay_mat, 2, function(x) wilcox.test(x[Group_f %in% c(
          #    groups_to_compare$group1[g], groups_to_compare$group2[g])] ~ Group_f[Group_f %in% c(
          #      groups_to_compare$group1[g], groups_to_compare$group2[g])], data = new_curr_df, conf.int = TRUE)$estimate)
          
          ###### first define function
          hl <- function(x, y = NULL) {
            if (is.null(y)) {
              walsh <- outer(x, x, "+") / 2
              median(walsh[lower.tri(walsh, diag = TRUE)], na.rm = TRUE)
            } else {
              median(outer(x, y, "-"), na.rm = TRUE)
            }
          }
          
          #### then generate the vector
          curr_est <- c()
          for(i in colnames(Assay_mat)){
            ### x vs y so positive would mean up in x
            x <- new_curr_df[which(new_curr_df$Group_f == group_levels[2]),i]
            y <- new_curr_df[which(new_curr_df$Group_f == group_levels[1]),i]
            ce <- hl(x, y)
            names(ce) <- i
            curr_est <- c(curr_est,ce)
          }
          
          
          #### combine to df
          
          curr_p_df <- data.frame(curr_p_raw, curr_p_adj, curr_est) %>%
            rownames_to_column() 
          colnames(curr_p_df) <- c("rowname",
                                   
                                   paste0("p_raw"),
                                   paste0("p_adj"),
                                   paste0("est"))

        
        
        #### add to stats list
        list_of_stats[[paste0(fig_name_to_use,"_MW_2group")]] <- curr_p_df


```


```{r AssociationWithTOA_prep___other}
 cor_method <- "spearman" ### "spearman", "pearson"
  corr_vsOther_vars <- c("TOA") ### specify which non-olink variables you would like to correlate with the olink variables

  corr_group_var_names <- c("ALUM","ALFQA") ### these are are as in the spreadsheet, specify which values of the grouping variable you'd like to run corrplots for
 corr_timepoints <- c("WkPost","Change")
  
  ### suffix to name for output tab
  short_name_to_use <- "vsTOA"
  
corr_vsOther_pval <- 0.05 ## pvalue for annotation with stars & trimming of corrplot


```


```{r AssociationWithOther, results='asis', fig.width=7, fig.height=6}
#### prep dataframe to collect R and p
cor_df <- data.frame("rowname"=assays_to_use)

##### separately for each group and timepoint
for(curr_time in corr_timepoints){
  
  # cat("### ",paste0(curr_time," {.tabset}"),"\n")
  # cat('\n')
  
          ## list to collect for alluvials
   list_for_alluvial <- list()

   for(curr_group in corr_group_var_names){
    
    # cat("#### ",paste0(curr_group," {.tabset}"),"\n")
    # cat('\n')
    # 
    ### wide with columns = assays/variables for corr, rows = animals
    curr_df_trim <- curr_phen %>%
      ### trim to only samples in this timepoint and group
      dplyr::filter(Time_Point %in% curr_time & group %in% curr_group) %>%
      ### select only the olink columns and the TAO 
      dplyr::select(all_of(intersect(c(assays_to_use,corr_vsOther_vars,"Sample_ID"), colnames(.)))) %>%
      
      filter_all(any_vars(!is.na(.))) %>% ### removes NA rows (unnecessary)
      select(where(function(x) !all(is.na(x)))) ### removes NA columns, as these might be specific per group
    

    
    ### set name
    set_name <- paste0(curr_group,"__",curr_time)
    
    ### run correlations - saves cor mat, raw p mat
    
     curr_columns_for_corr <- intersect(colnames(curr_df_trim),assays_to_use)
  curr_other_columns_for_corr <- intersect(corr_vsOther_vars,colnames(curr_df_trim))
  
  #### only if have other columns
  if(length(curr_other_columns_for_corr) > 0 & length(curr_columns_for_corr) > 0){
    
    
 ## generate cor matrix 
  if(length(corr_vsOther_vars) == 1){
    otherVar_df <- data.frame(as.numeric(data.frame(curr_df_trim)[,corr_vsOther_vars])) 
    colnames(otherVar_df) <- corr_vsOther_vars
      cor_mat <- cor(as.matrix(curr_df_trim[,curr_columns_for_corr]), as.matrix(otherVar_df), method = cor_method, use="pa")
  }else{
      cor_mat <- cor(as.matrix(data.frame(curr_df_trim)[,curr_columns_for_corr]), as.matrix(curr_df_trim[,curr_other_columns_for_corr]), method = cor_method, use="pa")
    
  }
  

  ## raw p
  cor_test_mat <- data.frame(matrix(NA, nrow=length(curr_columns_for_corr), ncol=length(curr_other_columns_for_corr)))
  rownames(cor_test_mat) <- curr_columns_for_corr
  colnames(cor_test_mat) <- curr_other_columns_for_corr
  
  for(curr1 in curr_columns_for_corr) {
    for(curr2 in intersect(corr_vsOther_vars,colnames(curr_df_trim))) {
      
      #### only if have non-NAs for at least 3 animals
      animals_withBoth <- curr_df_trim[,c(curr1,curr2)] %>%
        dplyr::filter(!is.na(!!rlang::sym(curr1)) & !is.na(!!rlang::sym(curr2)))
      
      if(dim(animals_withBoth)[1] > 0){
              cor_test_mat[curr1, curr2] <- 
        cor.test(as.matrix(data.frame(curr_df_trim)[,curr1]), 
                 as.matrix(as.numeric(data.frame(curr_df_trim)[,curr2])), 
                 method = cor_method, use="pa")$p.value

      }else{
              cor_test_mat[curr1, curr2] <- NA

      }
      
    }
  }
  
 ## adj p for multiple comparisons by BH over assays
  cor_test_adj_mat <- cor_test_mat #duplicate dataframe to replace values
  for(curr_assay in curr_columns_for_corr) {
    cor_test_adj_mat[curr_assay,] <- p.adjust(cor_test_mat[curr_assay,], method = "BH")
  }


 
    ### save cor_mat, cor_test_mat to combo df
    cor_df <- cor_df %>%
      dplyr::left_join(data.frame(cor_mat) %>%
                         ### rename to have the set name as column name
                         `colnames<-`(c(paste0(colnames(.),"_",set_name,"_R"))) %>%
                         rownames_to_column(),
                       by = "rowname") %>%
      dplyr::left_join(data.frame(cor_test_mat) %>%
                         ### rename to have the set name as column name
                         `colnames<-`(c(paste0(colnames(.),"_",set_name,"_p"))) %>%
                         rownames_to_column(),
                       by = "rowname")
    
    
    
    
    #### generate long to get number of sig pairs
    curr_Rp <- data.frame(cor_mat) %>%
      ### rename to have the set name as column name
      `colnames<-`(c(paste0(colnames(.),"_",set_name,"_R"))) %>%
      rownames_to_column() %>%
      dplyr::left_join(data.frame(cor_test_mat) %>%
                         ### rename to have the set name as column name
                         `colnames<-`(c(paste0(colnames(.),"_",set_name,"_p"))) %>%
                         rownames_to_column(),
                       by = "rowname") %>%
      tidyr::pivot_longer(!rowname, names_to = "assay_group_tp_stat", values_to = "value") %>% 
      dplyr::mutate(stat = str_after_last(assay_group_tp_stat,"_"), 
                    assay_group_tp = str_before_last(assay_group_tp_stat,"_"), 
                    assay_group_tp_stat = NULL) %>%
      tidyr::pivot_wider(names_from = "stat", values_from = "value") %>%
      ### remove self associations
      dplyr::filter(rowname != str_before_first(assay_group_tp,paste0("_",curr_group)))
    
    list_for_alluvial[[paste0(curr_group)]] <- curr_Rp  
    
    
    trim_pval_level <- corr_vsOther_pval
    
  }else{
    
  }
     
    
    
  }
  cat("\n")
  cat(' \n \n') ### this is the key!
  
  
  
 
    
    cat("\n")
    cat(' \n \n') ### this is the key!
    
}
  cat("\n")
    cat(' \n \n') ### this is the key!
 
    
    corVSTOA_df_prep <- cor_df %>% 
      tidyr::pivot_longer(!rowname, names_to = "assay_group_tp_stat", values_to = "value") %>% 
      dplyr::mutate(stat = str_after_last(assay_group_tp_stat,"_"), 
                    assay_group_tp = str_before_last(assay_group_tp_stat,"_"), 
                    assay_group = str_before_last(assay_group_tp,"__"),
                    tp1 = str_after_last(assay_group_tp,"__"),
                    assay2 = str_before_last(assay_group,"_"),
                    AnimalsFrom = str_after_last(assay_group,"_"),
                    assay1 = paste0(rowname,"__",tp1)) %>%
      dplyr::select(-c(assay_group_tp_stat,assay_group_tp,assay_group,tp1,rowname)) %>% 
      dplyr::relocate(assay1,assay2,AnimalsFrom) %>%
      tidyr::pivot_wider(names_from = "stat", values_from = "value") %>%
      ### remove rows without R and p
      dplyr::filter(!is.na(R)) %>%
  ### make wider
  tidyr::pivot_wider(names_from = "AnimalsFrom", values_from = c("R","p"))
    
    
    ######### save to stats if not 6a (since will get saved as supp 7a)
    if(fig_name_to_use != "Fig6a"){
      list_of_stats[[paste0(fig_name_to_use,"_",short_name_to_use)]]  <- corVSTOA_df_prep
    }
    
```


```{r AssociationVsSelf_prep}
corr_timepoints <- list("WkPostAndChange"=c("WkPost","Change"))
corr_pval_cut <- 0.05 ## pvalue for trimming corrplot no quotes
```

```{r AssociationVsSelf, results='asis', fig.width=9, fig.height=9}

######## setup
cor_method <- "spearman" ### "spearman", "pearson"

  corr_group_var_names <- c("ALUM","ALFQA") ### these are are as in the spreadsheet, specify which values of the grouping variable you'd like to run corrplots for
    
  ### suffix to name for output tab
  short_name_to_use <- "vsSelf"

  

#### prep dataframe to collect R and p
cor_combo_df <- data.frame()

#### prep df to  collect trim stats
trim_stats_df <- data.frame()



##### separately for each group and timepoint
for(curr_time in names(corr_timepoints)){
    
      cat("#### ",paste0(curr_time," {.tabset}"),"\n")
      cat('\n')

    
    
         ## list to collect for alluvials
   list_for_alluvial <- list()
 for(curr_group in corr_group_var_names){
   

   
      cat("##### ",paste0(curr_group," {.tabset}"),"\n")
      cat('\n')
      
      ### if both
      if(grepl("And",curr_time)){
           ### wide with columns = assays/variables for corr, rows = animals
      curr_df_trim <- curr_phen %>%
        ### trim to only samples in this timepoint and group
        dplyr::filter(Time_Point %in% corr_timepoints[[curr_time]] & group %in% curr_group) %>%
                ### select only the assays_to_use
        dplyr::select(all_of(intersect(c(assays_to_use,"Sample_ID","Time_Point"), colnames(.)))) %>%
                ### add tp to variable name
        ### first make longer
        tidyr::pivot_longer(cols = all_of(assays_to_use), names_to = "variable", values_to = "value") %>%
        ####### remove ADCP ADNP and trogo WkPost
        dplyr::mutate(helper = ifelse(grepl(paste(c("ADCP","ADNP","Trogo"),collapse = '|'),variable) & Time_Point == "WkPost",
                                      "cut","keep")) %>%
        dplyr::filter(helper == "keep") %>%
        dplyr::mutate(variable =paste0(variable,"_",Time_Point),
                      Time_Point = NULL,
                      helper = NULL) %>%
        ### make wider again
        tidyr::pivot_wider(names_from = "variable", values_from = "value") %>%


             filter_all(any_vars(!is.na(.))) %>% ### removes NA rows (unnecessary)
             select(where(function(x) !all(is.na(x)))) ### removes NA columns, as these might be specific per group
      

      }else{
      
      ### wide with columns = assays/variables for corr, rows = animals
      curr_df_trim <- curr_phen %>%
        ### trim to only samples in this timepoint and group
        dplyr::filter(Time_Point %in% corr_timepoints[[curr_time]] & group %in% curr_group) %>%
                ### select only the assays_to_use
        dplyr::select(all_of(intersect(c(assays_to_use,"Sample_ID"), colnames(.)))) %>%

             filter_all(any_vars(!is.na(.))) %>% ### removes NA rows (unnecessary)
             select(where(function(x) !all(is.na(x)))) ### removes NA columns, as these might be specific per group
      }


      ### set name
      set_name <- paste0(curr_group,"__",curr_time)
      curr_assays_to_use <- setdiff(colnames(curr_df_trim),c("Sample_ID","TOA"))
      
      ### run correlations - saves cor mat, raw p mat
     curr_columns_for_corr <- intersect(colnames(curr_df_trim),curr_assays_to_use)
 ## generate cor matrix 

      cor_mat <- cor(as.matrix(data.frame(curr_df_trim)[,curr_columns_for_corr]), method = cor_method, use="pa")
    

  ## raw p
  cor_test_mat <- data.frame(matrix(NA, nrow=length(curr_columns_for_corr), ncol=length(curr_columns_for_corr)))
  rownames(cor_test_mat) <- curr_columns_for_corr
  colnames(cor_test_mat) <- curr_columns_for_corr
  
  for(curr1 in curr_columns_for_corr) {
    for(curr2 in curr_columns_for_corr) {
      
      if(curr1 != curr2){
             #### only if have non-NAs for at least 3 animals
      animals_withBoth <- curr_df_trim[,c(curr1,curr2)] %>%
        dplyr::filter(!is.na(!!rlang::sym(curr1)) & !is.na(!!rlang::sym(curr2)))

      }else{
        ### for diagonal make dummy helper df
        animals_withBoth <- data.frame(matrix(nrow=2))
      }
 
      if(dim(animals_withBoth)[1] > 0){
              cor_test_mat[curr1, curr2] <- 
        cor.test(as.matrix(data.frame(curr_df_trim)[,curr1]), 
                 as.matrix(data.frame(curr_df_trim)[,curr2]), 
                 method = cor_method, use="pa")$p.value

      }else{
              cor_test_mat[curr1, curr2] <- NA

      }
      
    }
  }
  
 ## adj p for multiple comparisons by BH over assays
  cor_test_adj_mat <- cor_test_mat #duplicate dataframe to replace values
  for(curr_assay in curr_columns_for_corr) {
    cor_test_adj_mat[curr_assay,] <- p.adjust(cor_test_mat[curr_assay,], method = "BH")
  }



#### generate long to get number of sig pairs
  curr_Rp <- data.frame(cor_mat) %>%
                               ### rename to have the set name as column name
                               `colnames<-`(c(paste0(colnames(.),"_",set_name,"_R"))) %>%
                               rownames_to_column() %>%
    dplyr::left_join(data.frame(cor_test_mat) %>%
                               ### rename to have the set name as column name
                               `colnames<-`(c(paste0(colnames(.),"_",set_name,"_p"))) %>%
                               rownames_to_column(),
                     by = "rowname") %>%
    tidyr::pivot_longer(!rowname, names_to = "assay_group_tp_stat", values_to = "value") %>% 
      dplyr::mutate(stat = str_after_last(assay_group_tp_stat,"_"), 
                    assay_group_tp = str_before_last(assay_group_tp_stat,"_"), 
                    assay_group_tp_stat = NULL) %>%
    tidyr::pivot_wider(names_from = "stat", values_from = "value") %>%
    ### remove self associations
    dplyr::filter(rowname != str_before_first(assay_group_tp,paste0("_",curr_group)))
  
  list_for_alluvial[[paste0(curr_group)]] <- curr_Rp  
  

    ### save cor_mat, cor_test_mat to combo df
  cor_combo_df <- rbind(cor_combo_df,curr_Rp)

      
      trim_pval_level <- corr_pval_cut

      cat("\n")
      cat(' \n \n') ### this is the key!
      
      ### paste number of pairs - needs to be divided by 2 to remove the duplicates
            cat(" ",paste0(length(which(curr_Rp$p < trim_pval_level & curr_Rp$R > 0))/2,
                           " up and ",
                           length(which(curr_Rp$p < trim_pval_level & curr_Rp$R < 0))/2,
                           " down pairs with p < ",trim_pval_level),"\n")
      cat('\n')
         
      
      cat("\n")
      cat(' \n \n') ### this is the key!
      
      ########## print table trimmed
      
  
      
 to_print_prep <-  curr_Rp %>%
   ## trim to p cut
   dplyr::filter(p < trim_pval_level) %>%
   ### remove group & tp
   dplyr::mutate(assay2 = str_remove_all(assay_group_tp, paste0("_",curr_group,"__",curr_time)),
                 assay_group_tp = NULL) %>%
   dplyr::rename(assay1=rowname, spearman_R=R, spearman_p=p) %>%
   dplyr::relocate(assay1, assay2) %>%
   ### remove inverse dups
   group_by(grp = paste(pmax(assay1, assay2), pmin(assay1, assay2), sep = "_")) %>%
 dplyr::filter(row_number() == 1) %>%
 ungroup() %>%
 dplyr::select(-grp) %>%
   ## add time to olink
   dplyr::mutate(assay1 = ifelse(grepl("_",assay1),
                                 assay1,
                                 paste0(assay1,"_",curr_time)),
                 assay2 = ifelse(grepl("_",assay2),
                                 assay1,
                                 paste0(assay2,"_",curr_time))) %>%
   #### add MW dir for each assay
   dplyr::left_join(list_of_stats[[paste0(fig_name_to_use,"_MW_2group")]] %>%
                      ## remove __ and make Wk generic
                      dplyr::mutate(rowname = ifelse(grepl("Wk",rowname),
                                                     paste0(str_before_last(rowname,"Wk"),"WkPost"),
                                                     rowname),
                                    rowname = str_replace_all(rowname,"__","_"),
                                    assay1_MW_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(rowname,assay1_MW_dir) %>%
                      ### rename rowname to assay1
                      dplyr::rename(assay1=rowname),
                    by = "assay1") %>%
      dplyr::left_join(list_of_stats[[paste0(fig_name_to_use,"_MW_2group")]] %>%
                      ## remove __ and make Wk generic
                      dplyr::mutate(rowname = ifelse(grepl("Wk",rowname),
                                                     paste0(str_before_last(rowname,"Wk"),"WkPost"),
                                                     rowname),
                                    rowname = str_replace_all(rowname,"__","_"),
                                    assay2_MW_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(rowname,assay2_MW_dir) %>%
                      ### rename rowname to assay1
                      dplyr::rename(assay2=rowname),
                    by = "assay2") 
 
 ####### if have TOA for this group
 if(length(which(grepl(curr_group,colnames(list_of_stats[[paste0(fig_name_to_use,"_vsTOA")]])) == TRUE)) > 0) {
   to_print_prep2 <- to_print_prep %>%
   #### add vsTOA for each assay
      dplyr::left_join(list_of_stats[[paste0(fig_name_to_use,"_vsTOA")]] %>%
                         ### trim to only this group
                         dplyr::select(assay1,!!!rlang::syms(colnames(.)[grepl(curr_group,colnames(.))])) %>%
                         ### make colnames generic
                         `colnames<-`(c(str_remove_all(colnames(.),curr_group))) %>%
                      ## remove __ 
                      dplyr::mutate(assay1 = str_replace_all(assay1,"__","_"),
                                    assay1_vsTOA_dir = ifelse(p_ < 0.05,
                                                    ifelse(R_ > 0,
                                                           "pos",
                                                           "neg"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay1,assay1_vsTOA_dir),
                    by = "assay1") %>%
    dplyr::left_join(list_of_stats[[paste0(fig_name_to_use,"_vsTOA")]] %>%
                         ### trim to only this group
                         dplyr::select(assay1,!!!rlang::syms(colnames(.)[grepl(curr_group,colnames(.))])) %>%
                         ### make colnames generic
                         `colnames<-`(c(str_remove_all(colnames(.),curr_group))) %>%
                      ## remove __ 
                      dplyr::mutate(assay2 = str_replace_all(assay1,"__","_"),
                                    assay2_vsTOA_dir = ifelse(p_ < 0.05,
                                                    ifelse(R_ > 0,
                                                           "pos",
                                                           "neg"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay2,assay2_vsTOA_dir),
                    by = "assay2") 
   
   #### save to plotted
   trim_stats_df <- rbind(trim_stats_df, to_print_prep2 %>%
                            ### add group
                            dplyr::mutate(group = paste0(curr_group)) %>%
                            dplyr::relocate(group)
   )
   
   

   
  to_print <- to_print_prep2 %>%
   
     DT::datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           iDisplayLength = 50,
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
   print(htmltools::tagList(to_print))

        cat("\n")
      cat(' \n \n') ### this is the key!
      
 }else{
   
   to_print_prep2 <- to_print_prep %>%
     ### add dummy columns
     dplyr::mutate(assay1_vsTOA_dir = "not_available",
                   assay2_vsTOA_dir = "not_available")
     
   
    #### save to plotted
   trim_stats_df <- rbind(trim_stats_df, to_print_prep2 %>%
                            ### add group
                            dplyr::mutate(group = paste0(curr_group)) %>%
                            dplyr::relocate(group)
   )
   
  
   to_print <- to_print_prep %>%
      DT::datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           iDisplayLength = 50,
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
   print(htmltools::tagList(to_print))
 }
    
      
      
    }
    cat("\n")
    cat(' \n \n') ### this is the key!
    
    
        
   ########### now alluvial
    
   # for(curr_trim_level in c(0.05, 0.01, 0.001)){
     curr_trim_level <- corr_pval_cut
    
        cat("##### ",paste0(" alluvial (",curr_trim_level,") "),"\n")
      cat('\n')
  
        df_for_alluvial_prep <-  list_for_alluvial[["ALUM"]] %>%
                    ### add in the ALFQA
          dplyr::full_join(list_for_alluvial[["ALFQA"]], by = colnames(.)) %>%
          ### remove rows without R and p
      dplyr::filter(!is.na(R))  %>%
          ### separate off timepoint and group
          dplyr::mutate(assay2 = str_before_last(str_before_last(assay_group_tp,"__"),"_"),
                        group = str_after_last(str_before_last(assay_group_tp,"__"),"_"),
                        assay_group_tp = NULL) %>%
          dplyr::rename(assay1 = rowname) %>%
          dplyr::relocate(assay1, assay2, group) %>%
          ### make wider combining group/stat
          tidyr::pivot_longer(cols = c(R,p), names_to = "stat", values_to = "value") %>%
          dplyr::mutate(group_stat = paste0(group,"_",stat),
                        group = NULL,
                        stat = NULL) %>%
          tidyr::pivot_wider(names_from = "group_stat", values_from = "value")  %>%
          ### add dir
          dplyr::mutate(ALFQA_dir = ifelse(ALFQA_p < curr_trim_level,
                                            ifelse(ALFQA_R > 0, "pos","neg"),
                                           "not_sig"),
                        ALUM_dir = ifelse(is.na(ALUM_p),"no_test",
                                          ifelse(ALUM_p < curr_trim_level,
                                            ifelse(ALUM_R > 0, "pos","neg"),
                                           "not_sig"))) %>%
    ## remove inverse duplicates
 group_by(grp = paste(pmax(assay1, assay2), pmin(assay1, assay2), sep = "_")) %>%
 dplyr::filter(row_number() == 1) %>%
 ungroup() %>%
 dplyr::select(-grp) %>%
  ### remove not_sig in both, and no_test since can't compare ALFQA/ALUM
  dplyr::mutate(helper = paste0(ALFQA_dir,"_",ALUM_dir)) %>%
  dplyr::filter(helper != "not_sig_not_sig") %>%
  dplyr::filter(ALUM_dir != "no_test") %>%
  ### also remove unneeded columns
  dplyr::select(-c(contains("_R"),contains("_p"),helper)) %>%
  ##make longer
  tidyr::pivot_longer(cols = c(ALFQA_dir,ALUM_dir), names_to = "group", values_to = "dir") %>%
  dplyr::mutate(group = str_remove_all(group,"_dir"),
                assay_only1 = str_before_last(assay1,"_"),
                assay_only2 = str_before_last(assay2,"_"),
                assay_pair = paste0(assay1,"__vs__",assay2))
        
        #### for WkPost AND chang
        if(grepl("And",curr_time)){
          df_for_alluvial <- df_for_alluvial_prep %>%
              ### remove associations between wkpost and change of same assay
  dplyr::filter(assay_only1 != assay_only2) %>%
  ## factorize
  dplyr::mutate(assay_pair_f = factor(assay_pair),
                group_f = factor(group, c("ALFQA","ALUM")),
                dir_f = factor(dir, c("pos","not_sig","neg"))) %>%
  dplyr::arrange(group_f, dir_f, assay_pair_f)

                        

            
        }else{
           df_for_alluvial <- df_for_alluvial_prep %>%

  ## factorize
  dplyr::mutate(assay_pair_f = factor(assay_pair),
                group_f = factor(group, c("ALFQA","ALUM")),
                dir_f = factor(dir, c("pos","not_sig","neg"))) %>%
  dplyr::arrange(group_f, dir_f, assay_pair_f)

                        

        }
        
        


### set up colors
color_df <- data.frame("group"=unique(levels(df_for_alluvial$dir_f)),
                       "color"=c("purple","grey","orange"))



 
gg <- ggplot(df_for_alluvial,
             aes(x = group_f, stratum = dir_f, alluvium = assay_pair_f))+
  stat_alluvium(aes(fill = dir_f), width = 1/2, alpha = 1/2,) +
  stat_stratum(aes(stratum = dir_f, fill = dir_f), width = 1/2) +
  scale_fill_manual(values = color_df$color)+
  stat_alluvium(geom = "text", aes(label = assay_pair_f), color = "black", size = 1.5, show.legend = FALSE) + ## to label NA category
  theme_bw() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())
plot(gg)


cat("\n")
cat(' \n \n') ### this is the key!



    }
    cat("\n")
cat(' \n \n') ### this is the key!


###### save correlation stats
list_of_stats[[paste0(fig_name_to_use,"_",short_name_to_use)]] <- cor_combo_df %>%
  ### make longer
  tidyr::pivot_longer(!c(rowname,assay_group_tp), names_to = "stat", values_to = "value") %>%
  ## break apart assay group time stat
  dplyr::rename(assay1 = rowname) %>%
  dplyr::mutate(assay2_group = str_before_last(assay_group_tp,"__"),
                tp = str_after_last(assay_group_tp,"__"),
                assay2 = str_before_last(assay2_group,"_"),
                group_stat = paste0(str_after_last(assay2_group,"_"),"_",stat),
                assay_group_tp = NULL,
                assay2_group = NULL,
                stat = NULL) %>%
  ### make wider
  tidyr::pivot_wider(names_from = "group_stat", values_from = "value")


#### save to plotted

   list_of_plotted[[paste0(fig_name_to_use,"_trimStats")]] <- trim_stats_df %>%
     dplyr::arrange(group, assay1, assay2)

```






## Supp Figure 2 (Functional Assays in Blood) {.tabset}

```{r scatterVsGroupSetup___SuppFig2abcdefhi}
composite_fig_name <- "SuppFig2abcdefhi"
scatter_df <- data.frame("name"=c("SuppFig2a",
                                  "SuppFig2b",
                                  "SuppFig2c",
                                  "SuppFig2d",
                                  "SuppFig2e",
                                  "SuppFig2f",
                                  "SuppFig2h",
                                  "SuppFig2i"),
                         "y"=c("Plasma_gp120_antibody_titers__WkPost",
                               "Plasma_AvidityScore_dV1M766gp120__WkPost",
                               "Plasma_AvidityScore_cV2gp120__WkPost",
                               "Plasma_ADCC_Titer__WkPost",
                               "Plasma_ADCC_Killing__WkPost",
                               "Plasma_V2specific_ADCC_Killing__WkPost",
                               "Plasma_Trogocytosis_WTgp120__Change",
                               "Plasma_ADCP_SIVgp120WT__Change"),
                         "trans_y"=c("log10",
                                     "log10",
                                     "log10",
                                     "log10",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other"))

fill_colors_to_use <- c(ALFQA="red",Alum="blue")
outline_colors_to_use <- c("black","black")

```

```{r scatterplotsVsGroupPlot___SuppFig2abcdefhi, ref.label=c('scatterplotsVsGroupPlot'), results='asis', fig.height=3, fig.width=4}
```



```{r scatterVsOtherSetup___SuppFig2gjkl}
composite_fig_name <- "SuppFig2gjkl"

scatter_df <- data.frame("name"=c("SuppFig2g",
                                  "SuppFig2j",
                                  "SuppFig2k",
                                  "SuppFig2l"),
                         "x"=c("Plasma_ADCC_Killing__WkPost",
                               "Plasma_ADCP_SIVgp120dV1__Change",
                               "Plasma_Trogocytosis_dV1gp120__Change",
                               "Plasma_AvidityScore_dV1M766gp120__WkPost"),
                         "trans_x"=c("identity",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "y"=c("Plasma_gp120_antibody_titers__WkPost",
                               "Plasma_ADNP_SIVgp120dV1__Change",
                               "Plasma_ADNP_SIVgp120dV1__Change",
                               "Plasma_ADNP_SIVgp120dV1__Change"),
                         "trans_y"=c("log10",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other",
                                     "other",
                                     "other"))

```

```{r scatterplotsVsOtherPlot___SuppFig2gjkl, ref.label=c('scatterplotsVsOtherPlot'), results='asis', fig.height=4, fig.width=5}
```





## Figure 3 (RM) {.tabset}


```{r scatterVsGroupSetup___Fig3aeh}
composite_fig_name <- "Fig3aeh"
scatter_df <- data.frame("name"=c("Fig3a",
                                  "Fig3e",
                                  "Fig3h"),
                         "y"=c("RectMucosa_CD73posCD163posMacrophages__Change",
                               "RectMucosa_NKp44__Change",
                               "RectMucosa_NKp44negNKG2Aneg_PMA_IFNg__Change"),
                         "trans_y"=c("identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other",
                                     "other"))

fill_colors_to_use <- c(ALFQA="red",Alum="blue")
outline_colors_to_use <- c("black","black")

```

```{r scatterplotsVsGroupPlot___Fig3aeh, ref.label=c('scatterplotsVsGroupPlot'), results='asis', fig.height=3, fig.width=4}
```




```{r scatterVsOtherSetup___Fig3b}
composite_fig_name <- ""
scatter_df <- data.frame("name"=c("Fig3b"),
                         "x"=c("PBMCs_PctCD14_efferocytes__WkPost"),
                         "trans_x"=c("identity"),
                         "y"=c("RectMucosa_CD73posCD163posMacrophages__Change"),
                         "trans_y"=c("identity"),
                         "dataset"=c("other"))

```

```{r scatterplotsVsOtherPlot___Fig3b, ref.label=c('scatterplotsVsOtherPlot'), results='asis', fig.height=4, fig.width=5}
```




```{r scatterVsTOASetup___Fig3cfg}
composite_fig_name <- "Fig3cfg"
scatter_df <- data.frame("name"=c("Fig3c",
                                  "Fig3f",
                                  "Fig3g"),
                         "y"=c("RectMucosa_DC10__Change",
                               "RectMucosa_NKp44__Change",
                               "RectMucosa_NKp44_IL17_gp120_stimulated__Change"),
                         "trans_y"=c("identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other",
                                     "other"))

```

```{r scatterplotsVsTOAPlot___Fig3cfg, ref.label=c('scatterplotsVsTOAPlot'), results='asis', fig.height=4, fig.width=5}
```






```{r scatterVsGroupSetup___Fig3di}
composite_fig_name <- "Fig3di"
scatter_df <- data.frame("name"=c("Fig3d",
                                  "Fig3i"),
                         "y"=c("Pct_CD73pos_CD163pos_macrophages__Rectal_mucosa__week_13",
                               "Pct_NKp44pos_ILCs_Rectal_mucosa__week_13"),
                         "trans_y"=c("identity",
                                     "identity"),
                         "dataset"=c("otherStudy",
                                     "otherStudy"))


fill_colors_to_use <- c("white","white")
outline_colors_to_use <- c(ALFQA="red",Alum="blue")

```

```{r scatterplotsVsGroupPlot___Fig3di, ref.label=c('scatterplotsVsGroupPlot'), results='asis', fig.height=3, fig.width=4}
```



```{r setup___Fig3j}
fig_name_to_use <- "Fig3j"
assays_to_use <- sort(unique(data_long$variable)[grepl("RectMucosa", unique(data_long$variable))])
corr_timepoints <- list("WkPostAndChange"=c("WkPost","Change"))
corr_pval_cut <- 0.05 ## pvalue for trimming corrplot no quotes
```

### `r paste0(fig_name_to_use)` {.tabset}

```{r setSamples_other___Fig3j, ref.label=c('setSamples_other')}
```

```{r MW_2group_other___Fig3j, ref.label=c('MW_2group_other')}
```

```{r AssociationWithTOA_prep___other___Fig3j, ref.label=c('AssociationWithTOA_prep___other')}
```

```{r AssociationWithTOA___Fig3j, ref.label=c('AssociationWithOther'), results='asis', fig.width=7, fig.height=6}
```

```{r AssociationVsSelf_prep___Fig3j, ref.label=c('AssociationVsSelf_prep')}
```

```{r AssociationVsSelf___Fig3j, ref.label=c('AssociationVsSelf'), results='asis', fig.width=9, fig.height=9}
```




## Supp Figure 3 (FACS scheme)

FACS scheme





## Supp Figure 4 (RM) {.tabset}

```{r scatterVsGroupSetup___SuppFig4bdehi}
composite_fig_name <- "SuppFig4bdehi"
scatter_df <- data.frame("name"=c("SuppFig4b",
                                  "SuppFig4d",
                                  "SuppFig4e",
                                  "SuppFig4h",
                                  "SuppFig4i"),
                         "y"=c("RectMucosa_CD163posMacrophages__Change",
                               "RectMucosa_DC10__Change",
                               "RectMucosa_CD73posDC10__Change",
                               "RectMucosa_NKG2Apos__Change",
                               "RectMucosa_NKp44negNKG2Aneg__Change"),
                         "trans_y"=c("identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other",
                                     "other",
                                     "other",
                                     "other"))


fill_colors_to_use <- c(ALFQA="red",Alum="blue")
outline_colors_to_use <- c("black","black")

```

```{r scatterplotsVsGroupPlot___SuppFig4bdehi, ref.label=c('scatterplotsVsGroupPlot'), results='asis', fig.height=3, fig.width=4}
```



```{r scatterVsOtherSetup___SuppFig4cfklmno}
composite_fig_name <- "SuppFig4cfklmno"
scatter_df <- data.frame("name"=c("SuppFig4c",
                                  "SuppFig4f",
                                  "SuppFig4k",
                                  "SuppFig4l",
                                  "SuppFig4m",
                                  "SuppFig4n",
                                  "SuppFig4o"
                                  ),
                         "x"=c("RectMucosa_CD73posCD163posMacrophages__Change",
                           "RectMucosa_CD73posCD163posMacrophages__Change",
                           "RectMucosa_DC10__Change",
                           "RectMucosa_CD163posMacrophages__Change",
                           "RectMucosa_CD163posMacrophages__Change",
                           "RectMucosa_CD73posCD163posMacrophages__WkPost",
                           "Plasma_ADCC_Killing__WkPost"),
                         "trans_x"=c("identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "y"=c("Vaginal_secretions_dV1gp120_antibody_titers__WkPost",
                           "RectMucosa_CD73posDC10__Change",
                           "RectMucosa_NKp44__WkPost",
                           "RectMucosa_NKp44__Change",
                           "RectMucosa_NKp44negNKG2Aneg_PMA_IFNg__Change",
                           "RectMucosa_NKp44negNKG2Aneg_PMA_IFNg__WkPost",
                           "RectMucosa_NKp44__Change"
                           ),
                         "trans_y"=c("identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other",
                                     "other"))

```

```{r scatterplotsVsOtherPlot___SuppFig4cfklmno, ref.label=c('scatterplotsVsOtherPlot'), results='asis', fig.height=4, fig.width=5}
```




```{r scatterVsGroupSetup___SuppFig4j}
composite_fig_name <- "SuppFig4j"
scatter_df <- data.frame("name"=c("SuppFig4j"),
                         "y"=c("Pct_IFN_NKG2AnegNKp44neg_ILCs_PMA_stimulated__Rectal_mucosa__week_13"),
                         "trans_y"=c("identity"),
                         "dataset"=c("otherStudy"))


fill_colors_to_use <- c("white","white")
outline_colors_to_use <- c(ALFQA="red",Alum="blue")

```

```{r scatterplotsVsGroupPlot___SuppFig4j, ref.label=c('scatterplotsVsGroupPlot'), results='asis', fig.height=3, fig.width=4}
```




## Figure 4 (Plasma cytokines) {.tabset}

```{r setSamples_olink}
### select samples from desired studies ONLY
curr_samples <- phen_olink$animal_time_study[which(phen_olink$Study_number %in% c("VB034","VB026-B","VB042") &
                                                          ### also add the baseline (will be duplicates)
                                         phen_olink$Time_Point %in% c("Baseline","wk12_24h","wk13") &
                                         phen_olink$Protective_Vaccine %in% c("YES"))]


curr_phen_prep <- phen_olink[which(phen_olink$animal_time_study %in% curr_samples),] 

####### curate olink assays 
## remove if not in all datasets or only very low in one dataset
### check number of NA in each olink assay
NA_df <- data.frame("assay"=olink_assay_columns,
                    "num_na_ALFQA"=rep(NA, length(olink_assay_columns)),
                    "num_na_AlumVB034"=rep(NA, length(olink_assay_columns)),
                    "num_na_AlumVB026B"=rep(NA, length(olink_assay_columns)))
for(curr_assay in olink_assay_columns){
  
  NA_df$num_na_ALFQA[which(NA_df$assay == curr_assay)] <- length(which(is.na(curr_phen_prep[which(curr_phen_prep$Study_number == "VB042"),curr_assay])))
  
  NA_df$num_na_AlumVB034[which(NA_df$assay == curr_assay)] <- length(which(is.na(curr_phen_prep[which(curr_phen_prep$Study_number == "VB034"),curr_assay])))
  
    NA_df$num_na_AlumVB026B[which(NA_df$assay == curr_assay)] <- length(which(is.na(curr_phen_prep[which(curr_phen_prep$Study_number == "VB026-B"),curr_assay])))
}

### if no ALFQA cannot compare ---> need to be removed from everything
assays_missing_ALFQA <- NA_df$assay[which(NA_df$num_na_ALFQA == 36)]
### if no VB026-B, then could compare non-baseline only (and will be 12 ALFQA vs 17 Alum animals) ----> can include in group comparisons only for wk12_24, or ALFQA only tests for all timepoints, BUT some of the assays missing ALFQA are present in Alum and these were removed, so probably should remove these as well
assays_missing_AlumVB026Bonly <- NA_df$assay[which(NA_df$num_na_ALFQA == 0 & 
                                                     NA_df$num_na_AlumVB026B == 24)]


curr_phen <- curr_phen_prep %>%
  dplyr::select(-c(!!!rlang::syms(c(assays_missing_ALFQA,assays_missing_AlumVB026Bonly)))) %>%
  ### make useful group
  dplyr::mutate(group = ifelse(Study_number == "VB042",paste0("ALFQA"),paste0("ALUM")))


olink_assay_columns_new <- intersect(olink_assay_columns, colnames(curr_phen))

#### save for later 
curr_phen_olink <- curr_phen

```

```{r PCAsetup___Fig4ab}
###### set params
group_order <- c("ALUM","ALFQA")
group_colors <- c("blue","red")
num_loadings_per_PC <- 5
ellipse_prob <-  0.95
pca_scale_to_use <- TRUE

######## set timepoints & Fig names df
fig_names_df <- data.frame("name"=c("Fig4a","Fig4b"),
                           "tp"=c("wk12_24h","wk13"))

```

```{r PCA, results='asis', fig.height=4, fig.width=5}


#### for each timepoint of interest
for(curr_time in fig_names_df$tp){
  curr_fig_name <- fig_names_df$name[which(fig_names_df$tp == curr_time)]
    cat("### ",paste0(curr_fig_name," {.tabset}"),"\n")
  cat('\n')
 
  
      cat(" ",paste0("ellipse = ",paste0(ellipse_prob*100)," % normal probability"),"\n")
  cat('\n')
 
  
     cat(" ",paste0("top ",paste0(num_loadings_per_PC)," loadings per PC displayed"),"\n")
  cat('\n')
  
  # for(pca_scale_to_use in c(TRUE,FALSE)){
  #   
  #       cat("#### ",paste0("scale = ", pca_scale_to_use," {.tabset}"),"\n")
  # cat('\n')

    
    
    ## make true long format
    curr_phen_long1 <- curr_phen %>%
      ### trim to timepoint
      dplyr::filter(Time_Point == curr_time) %>%
      tidyr::pivot_longer(all_of(olink_assay_columns_new), names_to = "Assay", values_to = "value") %>%
      ### also add useful group
      dplyr::mutate(log10_value = log10(value),
                    log2_value = log2(value)) %>%
      dplyr::arrange(group)
    
    
    list_of_plotted[[paste0(curr_fig_name)]] <- curr_phen_long1[,c("animal_time_study","group","Assay","value","log10_value")]
    
    
    ### generate dataframe for PCA, columns are assays, rows are samples
    curr_cm_prep <- curr_phen_long1 %>%
      ### trim to only the olink columns and the animal
      dplyr::select(c(Sample_ID,group,Assay,log10_value)) %>%
      ### combine animal and group
      dplyr::mutate(group_animal = paste0(group,"_",Sample_ID),
                    #group = NULL,
                    Sample_ID = NULL) %>%
      ### make wider
      tidyr::pivot_wider(names_from = "Assay", values_from = "log10_value") 
    
    
    curr_cm <- curr_cm_prep %>%
      ### remove columns with all NAs for at least one group
      dplyr::select(group, curr_cm_prep %>%
                      group_by(group) %>%
                      dplyr::summarise(across(everything(), ~!all(is.na(.x)))) %>%
                      ungroup() %>%
                      select(-group) %>%
                      select(where(all)) %>%
                      names()) %>%
      ###remove group
      dplyr::select(-group) %>%
      ### make group_animal rownames
      column_to_rownames("group_animal") 
    
    
    
    #### factorize curr_meta
    curr_meta <- curr_phen_long1 %>%
      ### select only the metadata
      dplyr::select(c(Sample_ID,group)) %>%
      ### make unique
      unique() %>%
      ### combine animal and group
      dplyr::mutate(group_animal = paste0(group,"_",Sample_ID)) %>%
      ### make group_animal rownames
      column_to_rownames("group_animal") %>%
      dplyr::mutate(group_f = factor(group, group_order))
    
    
    #### print thenumber of animals
  cat(" ",paste0(dim(curr_meta[which(curr_meta$group == "ALUM"),])[1]," ALUM + ",
            dim(curr_meta[which(curr_meta$group == "ALFQA"),])[1]," ALFQA animals in PCA"),"\n")
  cat('\n')

    
    
    ### generate PCA
    pca <- prcomp(curr_cm, scale. = pca_scale_to_use)
    
    ### calculate % for each PC
    pc_pct_list <- list()
    for(i in 1:10){
      pc_pct_list[[paste0("PC",i)]] <- round(pca$sdev[i]^2/sum(pca$sdev^2)*100,2)
    }
    m.pc1 = round(pca$sdev[1]^2/sum(pca$sdev^2)*100,2)
    m.pc2 = round(pca$sdev[2]^2/sum(pca$sdev^2)*100,2)
    m.pc3 = round(pca$sdev[3]^2/sum(pca$sdev^2)*100,2)
    m.pc4 = round(pca$sdev[4]^2/sum(pca$sdev^2)*100,2)
    
    
    ####### only need PC1 vs PC2 as others are not useful
    pcs_pairs <- list(c("PC1","PC2"))
    pca_pairs_list <- list(c(1:2))
    # pcs_pairs <- list(c("PC1","PC2"),c("PC3","PC4"),c("PC5","PC6"),c("PC7","PC8"))
    # pca_pairs_list <- list(c(1:2),c(3:4),c(5:6),c(7:8))
    
    for(pca_pair in 1:length(pcs_pairs)){
      pcs_to_plot <- pcs_pairs[[pca_pair]]
      pca_pairs <- pca_pairs_list[[pca_pair]]
      
      xlab=paste0(pcs_to_plot[1]," (",pc_pct_list[[paste0(pcs_to_plot[1])]]," %)")
      ylab=paste0(pcs_to_plot[2]," (",pc_pct_list[[paste0(pcs_to_plot[2])]]," %)")
      
      # cat("##### ",paste0(pcs_to_plot[1], " vs ", pcs_to_plot[2]),"\n")
      # cat('\n')
      # 
      ### trim to number of loadings you want    
      pca_trim <- pca
      scores <- data.frame(pca_trim$x[, pca_pairs])
      scores[] <- lapply(scores, function(x) x / sqrt(sum((x - mean(x))^2))) 
      loadings <- as.data.frame(pca_trim$rotation)[pca_pairs]
      
      scale <- min(
        max(abs(scores[,pcs_to_plot[1]]))/max(abs(loadings[,pcs_to_plot[1]])),
        max(abs(scores[,pcs_to_plot[2]]))/max(abs(loadings[,pcs_to_plot[2]]))
      )*0.8
      
      ## add labels and positions to loadings_df
      loadings_df <- data.frame(loadings * scale) %>%
        rownames_to_column("var") 
      
      ##trim    
      top_rots <- c()
      for(curr_pc in pcs_to_plot){
        pca_trim_rot <- loadings_df
        ### add sort column
        pca_trim_rot$PC_to_sort <- abs(pca_trim_rot[,curr_pc])
        ### collect top (pos or neg)
        pca_trim_rot <- pca_trim_rot %>%
          arrange(desc(PC_to_sort)) 
        top_rots <- c(top_rots, pca_trim_rot$var[1:num_loadings_per_PC])
      }
      
      ### now trim to only the top loadings
      top_unique <- unique(top_rots)
      loadings_df2 <- loadings_df[loadings_df$var %in% top_unique,]
      
      
      
      ######## save scores as plotted PCA
       list_of_plotted[[paste0(curr_fig_name,"_PCA")]] <- scores %>%
         rownames_to_column() %>%
         ## break up group & Sample_ID
         dplyr::mutate(group = str_before_first(rowname,"_"),
                       Sample_ID = str_after_first(rowname,"_"),
                       rowname = NULL) %>%
         dplyr::relocate(group,Sample_ID)
      
      ### plot for all paramters:
      
      #### base plot
      p <- ggplot(scores, aes(x = !!rlang::sym(pcs_to_plot[1]), y = !!rlang::sym(pcs_to_plot[2]))) + 
        stat_ellipse(aes(fill = curr_meta$group_f), geom = "polygon", level = as.numeric(ellipse_prob), inherit.aes = TRUE, alpha = 0.4) +
        geom_point(aes(fill=curr_meta$group_f), shape = 21, size = 3, stroke = 0.7)+
        scale_fill_manual(name="Group",values=group_colors <- c("blue","red")) +  
        xlab(xlab)+ylab(ylab)+
        guides(fill=guide_legend(override.aes=list(shape=21)),color=guide_legend(override.aes=list(shape=21))) + ### change shape in legend to have fillable quality
        
        
        ### add loadings on top for final plot
        geom_segment(data = loadings_df2,
                     aes(x = 0, y = 0, xend = !!rlang::sym(pcs_to_plot[1]), yend = !!rlang::sym(pcs_to_plot[2])), 
                     color = "grey40", arrow = arrow(angle = 25, length = unit(4, "mm")))+
        geom_text_repel(data = loadings_df2, aes(x = !!rlang::sym(pcs_to_plot[1]), y = !!rlang::sym(pcs_to_plot[2]), label = var), size=3.5, color="grey25",
                        min.segment.length = 0, segment.color = 'grey60') +
        #coord_fixed() +
        theme_bw()+  
        theme(panel.grid.major = element_blank() ,
        panel.grid.minor = element_blank())
      
      plot(p)
      
      
      cat('\n')
      cat(' \n \n') ### this is the key! 
      
      ##### also print loadings dff
      
        to_print <- loadings_df %>% 
          dplyr::arrange(desc(PC1)) %>%
    DT::datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           iDisplayLength = 25,
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
            
            #) , height = "100%", width = "100%")
    print(htmltools::tagList(to_print))

      
      
 #   }
    
    cat('\n')
    cat(' \n \n') ### this is the key! 
    
  }
  
  cat('\n')
  cat(' \n \n') ### this is the key! 
  
  

}

cat('\n')
cat(' \n \n') ### this is the key! 



```


```{r volcanoSetup___Fig4cd}
fig_names_df <- data.frame("name"=c("Fig4c","Fig4d"),
                           "tp"=c("wk12_24h","wk13"))

fig_name_simp <- "Fig4cd"
```

```{r MW_2group_olink}
#Mann-Whitney/Wilcoxon test between groups at each timepoint
#### get data into true wide format
## rows = animals (+ group and any metadata variables that are intrinsic (i.e. don't change with time) to the animal)
## columns = variables + time
group_levels <- c("ALUM","ALFQA")

data_wide <- curr_phen %>%
  ### remove unused metadata
  dplyr::select(c(group,Time_Point,Sample_ID,c(!!!rlang::syms(c(olink_assay_columns_new))))) %>%
  ###  make long
  tidyr::pivot_longer(cols = all_of(olink_assay_columns_new), names_to = "variable", values_to = "value") %>%
  ### combine variable + time 
  dplyr::mutate(var_time = paste0(variable,"__",Time_Point),
                ## remove variable and time
                variable = NULL, 
                Time_Point = NULL) %>%
  ### make wider
  tidyr::pivot_wider(names_from = "var_time", values_from = "value") %>%
  ### factorize group
  dplyr::mutate(Group_f = factor(group, group_levels),
                ## also make rownames be animal
                rowname = Sample_ID) %>%
  column_to_rownames()

list_of_plotted[[paste0(fig_name_simp)]] <- data_wide %>%
  ## remove columns with all na
  select_if(~sum(!is.na(.)) > 0) %>%
  ### trim to only these timepoints
  dplyr::select(group, Sample_ID, !!!rlang::syms(colnames(.)[grepl(paste(fig_names_df$tp, collapse = '|'),colnames(.))])) %>%
  ### also remove rows if have no data
    ### remove rows with all missing 
  dplyr::filter_at(vars(setdiff(colnames(.),c("group","Sample_ID"))), any_vars(!is.na(.))) 


###### trim to only variables with some variation (i.e. if all animals == 0 then removed)
data_trim <- data_wide %>%
  dplyr::select(where(~n_distinct(.) > 1))


        ### trim to only the variables with data for both groups (removes if all NA in one group, even if there is data in the other group because will have nothing to compare to)
        trim_df1 <- data_trim[which(data_trim$Group_f == group_levels[1]),] #%>% dplyr::select(-Group_f)
        columns_with_noNAs1 <- colnames(trim_df1)[which(colSums(is.na(trim_df1)) == 0)]
        columns_with_someNAs1 <- colnames(trim_df1)[which(colSums(is.na(trim_df1)) > 0 & colSums(is.na(trim_df1)) < dim(trim_df1)[1])]
        columns_with_allNAs1 <- colnames(trim_df1)[which(colSums(is.na(trim_df1)) == dim(trim_df1)[1])]
   
        
        trim_df2 <- data_trim[which(data_trim$Group_f == group_levels[2]),] #%>% dplyr::select(-Group_f)
        columns_with_noNAs2 <- colnames(trim_df2)[which(colSums(is.na(trim_df2)) == 0)]
        columns_with_someNAs2 <- colnames(trim_df2)[which(colSums(is.na(trim_df2)) > 0 & colSums(is.na(trim_df2)) < dim(trim_df2)[1])]
        columns_with_allNAs2 <- colnames(trim_df2)[which(colSums(is.na(trim_df2)) == dim(trim_df2)[1])]



        
      new_curr_df <- data_trim[,intersect(c(columns_with_noNAs1,columns_with_someNAs1),c(columns_with_noNAs2,columns_with_someNAs2))] 
      
      
        ## generate matrix with only variables that you want to test (no character variables and nothing that has zero variation across all animals)
      Assay_mat <- data.matrix(new_curr_df %>%
                               dplyr::select(-c(Sample_ID,group,Group_f)))
      
      
          curr_p_raw <- apply(Assay_mat, 2, function(x) wilcox.test(x ~ Group_f, data = new_curr_df)$p.value)
          curr_p_adj <- p.adjust(curr_p_raw, method = 'BH') 
          # curr_est <- apply(new_Assay_mat, 2, function(x) wilcox.test(x[Group_f %in% c(
          #    groups_to_compare$group1[g], groups_to_compare$group2[g])] ~ Group_f[Group_f %in% c(
          #      groups_to_compare$group1[g], groups_to_compare$group2[g])], data = new_curr_df, conf.int = TRUE)$estimate)
          
          ###### first define function
          hl <- function(x, y = NULL) {
            if (is.null(y)) {
              walsh <- outer(x, x, "+") / 2
              median(walsh[lower.tri(walsh, diag = TRUE)], na.rm = TRUE)
            } else {
              median(outer(x, y, "-"), na.rm = TRUE)
            }
          }
          
          #### then generate the vector
          curr_est <- c()
          for(i in colnames(Assay_mat)){
            ### x vs y so positive would mean up in x
            x <- new_curr_df[which(new_curr_df$Group_f == group_levels[2]),i]
            y <- new_curr_df[which(new_curr_df$Group_f == group_levels[1]),i]
            ce <- hl(x, y)
            names(ce) <- i
            curr_est <- c(curr_est,ce)
          }
          
          
          #### combine to df
          
          curr_p_df <- data.frame(curr_p_raw, curr_p_adj, curr_est) %>%
            rownames_to_column() 
          colnames(curr_p_df) <- c("rowname",
                                   
                                   paste0("p_raw"),
                                   paste0("p_adj"),
                                   paste0("est"))

        
        
        #### add to stats list
        list_of_stats[[paste0(fig_name_simp,"_MW_2group")]] <- curr_p_df


```


```{r volcanos, results='asis', fig.height=5, fig.width=7}

pointSize <- 2
labSize <- 4.5

data_for_volc_prep <- list_of_stats[[paste0(fig_name_simp,"_MW_2group")]] %>%
  ### separate timepoint from variable
  dplyr::mutate(variable = str_before_first(rowname,"__"),
                Time_Point = str_after_first(rowname,"__"),
                ###add log10(est*100) (multiply by 100 to get out of the 0-1 range)
                log10_est100 = ifelse(est == 0, 0,
                                      ifelse(est < 0,
                                  -1*log10(abs(est*100)),
                                  log10(est*100))),
                ## add dir to color by
                dir = ifelse(p_raw < 0.05,
                             ifelse(est > 0, "ALFQA > ALUM", "ALFQA < ALUM"),
                             "not_sig"))


#### save this as plotted values
list_of_plotted[[paste0(fig_name_simp,"_MW_2group")]] <- data_for_volc_prep 


######## for each timepoint

for(curr_time in fig_names_df$tp){
  curr_fig_name <- fig_names_df$name[which(fig_names_df$tp == curr_time)]
    cat("### ",paste0(curr_fig_name," {.tabset}"),"\n")
  cat('\n')
 
  
      cat(" ",paste0("x-axis: log10(abs(MW estimate x100)) x sign of MW estimate"),"\n")
  cat('\n')
  
        cat(" ",paste0("MW estimate = median of outer differences"),"\n")
  cat('\n')
  
        cat(" ",paste0("outer differences = difference between all pairs of values in group1 - group2, e.g. for A = a1, a2 and B = b1, b2; then the outer differences would be: a1-b1, a2-b1, a1-b2, a2-b2. And the MW estimate is the median of these deltas. For MW estimate > 0, ALFQA > ALUM; and for MW estmate < 0, ALFQA < ALUM"),"\n")
  cat('\n')
  
        cat(" ",paste0("x100 = to shift all estimates > 1 (or < -1) so that the log10 does not transform values -1 < x < 1; thus for a MW estimate of 0.01 (the smallest magnitude estimate) --> log10(0.01 x100) = 0"),"\n")
  cat('\n')
 
  data_for_volc <- data_for_volc_prep %>%
    ### trim to timepoint
    dplyr::filter(Time_Point == curr_time) %>%
  ### 
  dplyr::mutate(MW = factor(dir, c("ALFQA > ALUM","ALFQA < ALUM","not_sig")),
                ### add label only for sig
                label_print = ifelse(dir != "not_sig",variable,""))

    color_df <- data.frame("category"=c("ALFQA > ALUM","ALFQA < ALUM","not_sig"),
                             "colors"=c( "red","blue","grey60"))
    
      colors_to_use <- color_df$colors[which(color_df$category %in% unique(data_for_volc$MW))]
      
      p <- ggplot(data_for_volc,
                  aes(log10_est100,-log10(p_raw),color=MW, fill=MW, label=label_print)) +
        geom_point(size = pointSize, shape = 21, alpha = 0.65) + 
        theme_bw()  +
        geom_vline(xintercept = c(0),linetype="dashed",color = "grey60") +
        geom_hline(yintercept = -log10(0.05),linetype="dashed",color = "grey60") +

        geom_text_repel(max.overlaps = 20, size = labSize, 
                        point.padding = 4,
                        show.legend = FALSE) + #this removes the 'a' from the legend
        scale_x_continuous(limits = c(-6,6), breaks = seq(-6,6,1))+
        scale_color_manual(values=colors_to_use)+
        scale_fill_manual(values=colors_to_use)+
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
              axis.text = element_text(size = 13)) +
        guides(color=guide_legend(override.aes = list(size=4), title = "MW p < 0.05"),fill=guide_legend(title = "MW p < 0.05"))+
        scale_size(guide = 'none') ### remove the size from legend

      
plot(p)

       cat("\n")
  cat(' \n \n') ### this is the key!

}

     cat("\n")
  cat(' \n \n') ### this is the key!

```




```{r alluvialSetup___Fig4e}
fig_name_simp2 <- "Fig4e"
```

### `r paste0(fig_name_simp2)`

```{r alluvial, results='asis', fig.width=6, fig.height=6}


####### baseline same assays, without baseline
baselineDiff_vars <- list_of_plotted[[paste0(fig_name_simp,"_MW_2group")]]$variable[which(list_of_plotted[[paste0(fig_name_simp,"_MW_2group")]]$Time_Point == "Baseline" & list_of_plotted[[paste0(fig_name_simp,"_MW_2group")]]$dir != "not_sig")]


  cat(" ",paste0("baseline same assays, without Baseline"),"\n")
  cat('\n')


df_for_alluvial <- list_of_plotted[[paste0(fig_name_simp,"_MW_2group")]] %>%
  ### remove baseline for now
  dplyr::filter(Time_Point != "Baseline") %>%
  dplyr::filter(!variable %in% baselineDiff_vars) %>%
  ## only needed columns 
  dplyr::select(variable,Time_Point,dir) %>%
  ## remove variables not sig at both tiimepoints
  ## first pivot wider
  tidyr::pivot_wider(names_from = "Time_Point", values_from = "dir") %>%
  ## add helpter column
  dplyr::mutate(helper = paste0(wk12_24h,"_",wk13)) %>%
  dplyr::filter(helper != "not_sig_not_sig") %>%
  ## return to longer
  dplyr::select(-helper) %>%
  tidyr::pivot_longer(!variable, names_to = "Time_Point", values_to = "dir") %>%
  ## factorize
  dplyr::mutate(variable_f = factor(variable),
                Time_Point_f = factor(Time_Point,c("wk12_24h","wk13")),
                dir_f = factor(dir, c("ALFQA > ALUM","not_sig","ALFQA < ALUM"))) %>%
  dplyr::arrange(Time_Point_f, dir_f, variable_f)
  


### set up colors
color_df <- data.frame("group"=unique(levels(df_for_alluvial$dir_f)),
                       "color"=c("red","grey","blue"))



 
gg <- ggplot(df_for_alluvial,
             aes(x = Time_Point_f, stratum = dir_f, alluvium = variable_f))+
  stat_alluvium(aes(fill = dir_f), width = 1/2, alpha = 1/2,) +
  stat_stratum(aes(stratum = dir_f, fill = dir_f), width = 1/2) +
  scale_fill_manual(values = color_df$color)+
  stat_alluvium(geom = "text", aes(label = variable_f), color = "white", size = 3, show.legend = FALSE) + ## to label NA category
  theme_bw() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())
plot(gg)


cat("\n")
cat(' \n \n') ### this is the key!
list_of_plotted[[paste0(fig_name_simp2)]] <- list_of_plotted[[paste0(fig_name_simp)]]
list_of_plotted[[paste0(fig_name_simp2,"_MW_2group")]] <- df_for_alluvial[,c("variable","Time_Point","dir")]

```


```{r scatterVsGroupSetup___Fig4fghijk}
composite_fig_name <- "Fig4fghijk"
scatter_df <- data.frame("name"=c("Fig4f",
                                  "Fig4g",
                                  "Fig4h",
                                  "Fig4i",
                                  "Fig4j",
                                  "Fig4k"),
                         "y"=c("CCL2__wk12_24h",
                               "IL18__wk12_24h",
                               "IL18__wk13",
                               "CCL2__wk13",
                               "LTA__wk12_24h",
                               "LTA__wk13"),
                         "trans_y"=c("identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("olink",
                                     "olink",
                                     "olink",
                                     "olink",
                                     "olink",
                                     "olink"))

fill_colors_to_use <- c(ALFQA="red",Alum="blue")
outline_colors_to_use <- c("black","black")

```

```{r scatterplotsVsGroupPlot___Fig4fghijk, ref.label=c('scatterplotsVsGroupPlot'), results='asis', fig.height=3, fig.width=4}
```







## Supp Figure 5 (Plasma cytokines) {.tabset}

```{r setSamples_PCA___SuppFig5a, ref.label=c('setSamples_olink')}
```

```{r PCAsetup___SuppFig5a}
###### set params
group_order <- c("ALUM","ALFQA")
group_colors <- c("blue","red")
num_loadings_per_PC <- 5
ellipse_prob <-  0.95
pca_scale_to_use <- TRUE

######## set timepoints & Fig names df
fig_names_df <- data.frame("name"=c("SuppFig5a"),
                           "tp"=c("Baseline"))

```

```{r PCA___SuppFig5a, ref.label=c('PCA'), results='asis', fig.height=4, fig.width=5}
```


```{r volcanoSetup___SuppFig5b}
fig_names_df <- data.frame("name"=c("SuppFig5b"),
                           "tp"=c("Baseline"))

fig_name_simp <- "SuppFig5b"
```

```{r MW_2group_olink___SuppFig5b, ref.label=c('MW_2group_olink')}
```


```{r volcanos___SuppFig5b, ref.label=c('volcanos'), results='asis', fig.height=5, fig.width=7}
```



```{r alluvialSetup___SuppFig5c}
fig_name_simp2 <- "SuppFig5c"
```

### `r paste0(fig_name_simp2)`

```{r alluvialShortlist, results='asis', fig.width=6, fig.height=3}
#### finally only the baseline different vars

  cat(" ",paste0("Baseline diff only"),"\n")
  cat('\n')

df_for_alluvial <- list_of_stats[[paste0(fig_name_simp,"_MW_2group")]] %>%
  ### separate timepoint from variable
  dplyr::mutate(variable = str_before_first(rowname,"__"),
                Time_Point = str_after_first(rowname,"__"),
                ###add log10(est*100) (multiply by 100 to get out of the 0-1 range)
                log10_est100 = ifelse(est == 0, 0,
                                      ifelse(est < 0,
                                  -1*log10(abs(est*100)),
                                  log10(est*100))),
                ## add dir to color by
                dir = ifelse(p_raw < 0.05,
                             ifelse(est > 0, "ALFQA > ALUM", "ALFQA < ALUM"),
                             "not_sig")) %>%
  ## only vars different at baseline
  dplyr::filter(variable %in% baselineDiff_vars) %>%
  ## only needed columns 
  dplyr::select(variable,Time_Point,dir) %>%
  ## remove variables not sig at both tiimepoints
  ## first pivot wider
  tidyr::pivot_wider(names_from = "Time_Point", values_from = "dir") %>%
  ## add helpter column
  dplyr::mutate(helper = paste0(Baseline,"_",wk12_24h,"_",wk13)) %>%
  dplyr::filter(helper != "not_sig_not_sig_not_sig") %>%
  ## return to longer
  dplyr::select(-helper) %>%
  tidyr::pivot_longer(!variable, names_to = "Time_Point", values_to = "dir") %>%
  ## factorize
  dplyr::mutate(variable_f = factor(variable),
                Time_Point_f = factor(Time_Point,c("Baseline","wk12_24h","wk13")),
                dir_f = factor(dir, c("ALFQA > ALUM","not_sig","ALFQA < ALUM"))) %>%
  dplyr::arrange(Time_Point_f, dir_f, variable_f)
  


### set up colors
color_df <- data.frame("group"=unique(levels(df_for_alluvial$dir_f)),
                       "color"=c("red","grey","blue"))



 
gg <- ggplot(df_for_alluvial,
             aes(x = Time_Point_f, stratum = dir_f, alluvium = variable_f))+
  stat_alluvium(aes(fill = dir_f), width = 1/2, alpha = 1/2,) +
  stat_stratum(aes(stratum = dir_f, fill = dir_f), width = 1/2) +
  scale_fill_manual(values = color_df$color)+
  stat_alluvium(geom = "text", aes(label = variable_f), color = "white", size = 3, show.legend = FALSE) + ## to label NA category
  theme_bw() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())
plot(gg)



cat("\n")
cat(' \n \n') ### this is the key!
list_of_plotted[[paste0(fig_name_simp2)]] <- list_of_plotted[[paste0(fig_name_simp)]]
list_of_plotted[[paste0(fig_name_simp2,"_MW_2group")]] <- df_for_alluvial

```




```{r setSamples_withBaseline___SuppFig5de}

fig_names_df <- data.frame("name"=c("SuppFig5d","SuppFig5e"),
                           "tp"=c("wk12_24h","wk13"))

fig_name_simp <- "SuppFig5de"

### select samples from desired studies & all baselines
curr_samples <- c(phen_olink$animal_time_study[which(phen_olink$Study_number %in% c("VB034","VB026-B","VB042") &
                                                          ### also add the baseline (will be duplicates)
                                         phen_olink$Time_Point %in% c("Baseline","wk12_24h","wk13") &
                                         phen_olink$Protective_Vaccine %in% c("YES"))],
                  phen_olink$animal_time_study[which(phen_olink$Time_Point == "Baseline")]
)


curr_phen <- phen_olink[which(phen_olink$animal_time_study %in% curr_samples),] 
```


```{r vsBaselinePlots, results='asis', fig.height=9, fig.width=9}

  #### for each timepoint of interest
for(curr_time in fig_names_df$tp){
  curr_fig_name <- fig_names_df$name[which(fig_names_df$tp == curr_time)]
    cat("### ",paste0(curr_fig_name," {.tabset}"),"\n")
  cat('\n')
 
  ## make true long format
  curr_phen_long1 <- curr_phen %>%
    tidyr::pivot_longer(all_of(olink_assay_columns_new), names_to = "Assay", values_to = "value") %>%
    ### also add useful group
    dplyr::mutate(group = ifelse(Time_Point == "Baseline",paste0("Baseline"),
                                 ifelse(Study_number == "VB042",paste0("ALFQA"),paste0("ALUM")))) %>%
    ### now add in baseline for the studies of interest
    dplyr::full_join(curr_phen[which(curr_phen$animal_time_study %in% curr_samples[duplicated(curr_samples)]),] %>%
                       tidyr::pivot_longer(all_of(olink_assay_columns_new), names_to = "Assay", values_to = "value") %>%
                       ### also add useful group
                       dplyr::mutate(group = ifelse(Study_number == "VB042",
                                                    paste0("ALFQA_baseline"),
                                                    paste0("ALUM_baseline"))), by = c(colnames(.))) %>%
    dplyr::mutate(group = factor(group, c("ALFQA","ALFQA_baseline","Baseline","ALUM_baseline","ALUM")),
                  ## add log
                  log10_value = log10(value),
                  log2_value = log2(value)) %>%
    ### trim to only baseline that the curr_time
    dplyr::filter(Time_Point %in% c("Baseline",curr_time)) %>%
    dplyr::arrange(group)
  
  
  
      
    list_of_plotted[[paste0(curr_fig_name)]] <- curr_phen_long1[,c("animal_time_study","group","Assay","value","log10_value")]
 #  list_of_plotted[[paste0(curr_fig_name,"_dataWithB_",curr_time)]] <- curr_phen_long1[,c("animal_time_study","group","Assay","value","log10_value")]

### collect means for each Assay/group
df_means1 <- curr_phen_long1 %>%
  group_by(Assay, group) %>%
  dplyr::summarise(mean = mean(log10_value, na.rm=TRUE))

### remove assays that don't have both treatment groups for
curr_phen_long <- curr_phen_long1 %>%
  dplyr::filter(Assay %in% setdiff(unique(curr_phen_long1$Assay),unique(df_means1$Assay[which(is.nan(df_means1$mean))]))) %>%
  dplyr::mutate(Assay = factor(Assay))

df_means <- df_means1 %>%
  dplyr::filter(Assay %in% setdiff(unique(curr_phen_long1$Assay),unique(df_means1$Assay[which(is.nan(df_means1$mean))]))) %>%
  dplyr::mutate(Assay = factor(Assay))

### get direction
df_dir <- df_means %>%
  tidyr::pivot_wider(names_from = "group", values_from = "mean") %>%
  dplyr::mutate(delta = ALFQA - ALUM,
                MW_pvalue = factor(ifelse(ALFQA > ALUM, "ALFQA > ALUM","ALFQA < ALUM"),
                                   c("ALFQA > ALUM","ALFQA < ALUM")))


df_dirB <- df_means %>%
  tidyr::pivot_wider(names_from = "group", values_from = "mean") %>%
  dplyr::mutate(delta = ALFQA_baseline - ALUM_baseline,
                MW_pvalue = factor(ifelse(ALFQA_baseline > ALUM_baseline,
                                          "ALFQA_baseline > ALUM_baseline",
                                          "ALFQA_baseline < ALUM_baseline"),
                                   c("ALFQA_baseline > ALUM_baseline","ALFQA_baseline < ALUM_baseline")))

### get Assays sorted by baseline mean
Assay_order_byBmeans <- df_dir %>%
  dplyr::arrange(Baseline) %>%
  dplyr::select(Assay)

#### get stats
df_stats1 <- curr_phen_long %>%
  dplyr::mutate(Assay = factor(Assay, as.character(Assay_order_byBmeans$Assay))) %>%
  group_by(Assay) %>%
  wilcox_test(log10_value ~ group) %>%
  add_significance("p") %>%
  dplyr::filter(group1 == "ALFQA" & group2 == "ALUM") %>%
  add_xy_position(x = "Assay", dodge = 0) %>%
  ## add dir color
  dplyr::left_join(df_dir[,c("Assay","MW_pvalue")], by ="Assay")

df_stats <- df_stats1 %>%
  ### modify y position
  dplyr::mutate(y.position = -6,
                p.signif = ifelse(p.signif == "ns","","\U2217"))



#### get baseline stats
df_statsB1 <- curr_phen_long %>%
  dplyr::mutate(Assay = factor(Assay, as.character(Assay_order_byBmeans$Assay))) %>%
  group_by(Assay) %>%
  wilcox_test(log10_value ~ group) %>%
  add_significance("p") %>%
  dplyr::filter(group1 == "ALFQA_baseline" & group2 == "ALUM_baseline") %>%
  add_xy_position(x = "Assay", dodge = 0) %>%
  ## add dir color
  dplyr::left_join(df_dirB[,c("Assay","MW_pvalue")], by ="Assay")

df_statsB <- df_statsB1 %>%
  ### modify y position
  dplyr::mutate(y.position = -5.7,
                p.signif = ifelse(p.signif == "ns","","b"))
  

########## collect stats to save
list_of_stats[[paste0(curr_fig_name,"_MWwB_",curr_time)]] <- df_stats1 %>%
  ### remove unneeded columns 
  dplyr::select(-c(.y.,y.position,groups,x,xmin,xmax)) %>%
  ### add column for timepoint
  dplyr::mutate(timepoint = paste0(curr_time)) %>%
  ### add baseline stats
  dplyr::full_join(df_statsB1 %>%
                     ### remove unneeded columns 
                     dplyr::select(-c(.y.,y.position,groups,x,xmin,xmax)) %>%
                     ### add column for timepoint
                     dplyr::mutate(timepoint = "baseline"),
                   by = colnames(.)) %>%
  ### arrange by assay
  dplyr::arrange(Assay, rev(timepoint))


########### collect plotted means
list_of_plotted[[paste0(curr_fig_name,"_MeanswB_",curr_time)]] <- df_means %>%
  ### arrange by assay
  dplyr::arrange(Assay)




  ########### now sort by up/down, then baseline mean
  
  Assay_order_byUPDown <- df_dir %>%
    ### add p vals
    dplyr::left_join(df_stats[,c("Assay","p")], by = "Assay") %>%
    dplyr::mutate(sig_dir = factor(ifelse(p < 0.05,as.character(MW_pvalue),"not_sig"),c("ALFQA > ALUM","not_sig","ALFQA < ALUM")),
                  ### sort so most extreme on edges
                  p_with_dir = ifelse(MW_pvalue == "ALFQA > ALUM",p,
                                      ifelse(MW_pvalue == "ALFQA < ALUM",-p,p))) %>%
  dplyr::arrange(sig_dir,desc(Baseline)) %>%
  dplyr::select(Assay)
  
  #### get stats
df_stats1 <- curr_phen_long %>%
  dplyr::mutate(Assay = factor(Assay, rev(as.character(Assay_order_byUPDown$Assay)))) %>%
  group_by(Assay) %>%
  wilcox_test(log10_value ~ group) %>%
  add_significance("p") %>%
  dplyr::filter(group1 == "ALFQA" & group2 == "ALUM") %>%
  add_xy_position(x = "Assay", dodge = 0) %>%
  ## add dir color
  dplyr::left_join(df_dir[,c("Assay","MW_pvalue")], by ="Assay")

df_stats <- df_stats1 %>%
  ### modify y position
  dplyr::mutate(y.position = -6,
                p.signif = ifelse(p.signif == "ns","","\U2217"))
  

  #### get baseline stats
df_statsB1 <- curr_phen_long %>%
  dplyr::mutate(Assay = factor(Assay, rev(as.character(Assay_order_byUPDown$Assay)))) %>%
  group_by(Assay) %>%
  wilcox_test(log10_value ~ group) %>%
  add_significance("p") %>%
  dplyr::filter(group1 == "ALFQA_baseline" & group2 == "ALUM_baseline") %>%
  add_xy_position(x = "Assay", dodge = 0) %>%
  ## add dir color
  dplyr::left_join(df_dirB[,c("Assay","MW_pvalue")], by ="Assay")

df_statsB <- df_statsB1 %>%
  ### modify y position
  dplyr::mutate(y.position = -5.7,
                p.signif = ifelse(p.signif == "ns","","b"))

  
######### plot
### design plot with assay on x and value on y for stats to work
p <- ggplot(curr_phen_long[which(curr_phen_long$group == "Baseline"),] %>%
              dplyr::mutate(Assay = factor(Assay, rev(as.character(Assay_order_byUPDown$Assay)))), aes(x = Assay, y = log10_value))+
  geom_boxplot(color = "grey60", outlier.size = 0)+

    geom_beeswarm(data = df_means %>%
              dplyr::mutate(Assay = factor(Assay, rev(as.character(Assay_order_byUPDown$Assay)))) , aes(x=Assay, y=mean, fill = group, color = group, shape = group), dodge.width = 0.5)+
  scale_fill_manual(values = c("#FF0000","white","grey60","white","#0244FF"))+
  scale_color_manual(values = c('ALFQA > ALUM'="#FF0000",
                                'ALFQA < ALUM'="#0244FF",
                                'ALFQA_baseline > ALUM_baseline'="#FF0000",
                                'ALFQA_baseline < ALUM_baseline'="#0244FF",
                                "ALFQA"="#FF0000",
                                "ALFQA_baseline"="#FF0000",
                                "Baseline"="black",
                                "ALUM_baseline"="#0244FF",
                                "ALUM"="#0244FF")
                     ,
                     ### only have the stars in the legend
                     breaks = c('ALFQA > ALUM', 'ALFQA < ALUM',
                                'ALFQA_baseline > ALUM_baseline', 
                                'ALFQA_baseline < ALUM_baseline')
                     )+
  scale_shape_manual(values = c(21,21,24,21,21))+

  scale_y_continuous(limits = c(-6,4))+
  stat_pvalue_manual(data=df_stats[which(df_stats$group1 == "ALFQA" & 
                                         df_stats$group2 == "ALUM"),] ,
                     label = "p.signif", size = 4, color = "MW_pvalue",
                     ### remove bracket and hide non-sig
                     remove.bracket = TRUE, #hide.ns = TRUE,
                     family="Arial Unicode MS",
                     vjust = 0.45, ### to get vertically aligned with Assay label
                     coord.flip = TRUE)+
  
    #### baseline p-values
    stat_pvalue_manual(data=df_statsB[which(df_statsB$group1 == "ALFQA_baseline" & 
                                         df_statsB$group2 == "ALUM_baseline"),] ,
                     label = "p.signif", size = 4, color = "MW_pvalue",
                     ### remove bracket and hide non-sig
                     remove.bracket = TRUE, #hide.ns = TRUE,
                     family="Arial Unicode MS",
                     vjust = 0.45, ### to get vertically aligned with Assay label
                     coord.flip = TRUE)+

   ### then flip
coord_flip()+
  theme_bw()+

      guides(
        ### asterisk legend
        color=guide_legend(title = "MW p < 0.05",
                             override.aes=list(label=c("\U2217","\U2217",
                                                       "b","b"))),
        ### group legend
        fill= guide_legend(override.aes = list(color=c("#FF0000","#FF0000","black","#0244FF","#0244FF"))))+
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank())



 
 plot(p)
 
      cat("\n")
  cat(' \n \n') ### this is the key!


  
  
  
}


     cat("\n")
  cat(' \n \n') ### this is the key!


```








## Figure 5 (Plasma cytokines) {.tabset}



```{r setup___Fig5a}
fig_name_to_use <- "Fig5a"
assays_to_use <- olink_assay_columns_new
fig_name_simp <- "Fig5a"

#### dummy 
fig_names_df <- data.frame("tp"=c("Baseline","wk12_24h","wk13"))
```

### `r paste0(fig_name_to_use)` {.tabset}

```{r setSamples_corrVsSelf___Fig5a, ref.label=c('setSamples_olink')}
```

```{r MW_2group_olink___5a, ref.label=c('MW_2group_olink')}
```


```{r AssociationWithTOA_prep___olink}
 cor_method <- "spearman" ### "spearman", "pearson"
  corr_vsOther_vars <- c("TOA") ### specify which non-olink variables you would like to correlate with the olink variables

  corr_group_var_names <- c("ALUM","ALFQA") ### these are are as in the spreadsheet, specify which values of the grouping variable you'd like to run corrplots for
 corr_timepoints <- c("Baseline","wk12_24h","wk13")
  
  ### suffix to name for output tab
  short_name_to_use <- "vsTOA"
  
corr_vsOther_pval <- 0.05 ## pvalue for annotation with stars & trimming of corrplot


```


```{r AssociationWithTOA___Fig5a, ref.label=c('AssociationWithOther'), results='asis', fig.width=7, fig.height=6}
```


```{r AssociationVsSelf_prep___Fig5a}
  corr_timepoints <- list("Baseline"=c("Baseline"),
                          "wk12_24h"=c("wk12_24h"),
                          "wk13"=c("wk13"))
  corr_pval_cut <- 0.001 ## pvalue for trimming corrplot no quotes
```

```{r AssociationVsSelf___Fig5a, ref.label=c('AssociationVsSelf'), results='asis', fig.width=9, fig.height=9}
```



```{r Fig5bc_prep}
fig_names_df <- data.frame("name"=c("Fig5b","Fig5c"),
                          "tp"=c("wk12_24h","wk13"))

```

```{r IPA_inputTable, results='asis'}


#### for each timepoint
for(curr_time in fig_names_df$tp){
  
   curr_fig_name <- fig_names_df$name[which(fig_names_df$tp == curr_time)]

       cat("### ",paste0(curr_fig_name," "),"\n")
  cat('\n')
      cat(" ",paste0("For IPA (see methods) "),"\n")
  cat('\n')


to_print_prep <- curr_phen_olink %>%
  ### make longer
  tidyr::pivot_longer(cols = !c(animal_time_study,Study_number,Sample_ID,Time_Point,TOA,Protective_Vaccine,group), names_to = "assay", values_to = "value") %>%
  ### trim to only this time point
  dplyr::filter(Time_Point == curr_time) %>%
  ### add _mean to groupnames
  dplyr::mutate(group = paste0(group,"_mean")) %>%
  ### get ratio of means
  group_by(group,assay) %>%
  dplyr::summarise(mean = mean(value, na.rm = TRUE)) %>%
  ## make wider
  tidyr::pivot_wider(names_from = "group", values_from = "mean") %>%
  ## add FC ratio of means
  dplyr::mutate(FC_of_means = ALFQA_mean/ALUM_mean) %>%
  ### add in the MW stats
  dplyr::left_join(list_of_stats[["Fig5a_MW_2group"]] %>%
                     ##separate timepoint & assay
                     dplyr::mutate(assay = str_before_first(rowname,"__"),
                                   Time_Point = str_after_first(rowname,"__")) %>%
                     ### trim to curr_time
                     dplyr::filter(Time_Point == curr_time) %>%
                     ### remove unneeded columns
                     dplyr::select(-est, -Time_Point, -rowname, -p_adj),
                   by = "assay") %>%
  ### add in uniprot IDs
  dplyr::left_join(masterlist[["uniprotIDs"]], by = "assay") %>%
  dplyr::relocate(Uniprot, .after = "assay")
  

padj_vec <- p.adjust(to_print_prep$p_raw, method = "BH")
to_print_prep2 <- cbind(to_print_prep,data.frame("p_adj"=padj_vec))


to_print <- to_print_prep2 %>%
     DT::datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           iDisplayLength = 50,
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
   print(htmltools::tagList(to_print))

   
   
   #### also save to stats & plotted
   list_of_plotted[[paste0(curr_fig_name)]] <- curr_phen_olink %>%
     ### trim to only this time point
     dplyr::filter(Time_Point == curr_time)
   list_of_plotted[[paste0(curr_fig_name,"_IPAinput")]] <- to_print_prep2
   list_of_stats[[paste0(curr_fig_name)]] <- to_print_prep2


     cat("\n")
  cat(' \n \n') ### this is the key!

}
     cat("\n")
  cat(' \n \n') ### this is the key!


```


## Supp Figure 6 (Pathway analysis) {.tabset}

```{r SuppFig6_prep}
fig_names_df <- data.frame("name"=c("SuppFig6a","SuppFig6b"),
                          "tp"=c("wk12_24h","wk13"))

```

```{r IPA_inputTable___SuppFig6, ref.label=c('IPA_inputTable'), results='asis'}
```


## Figure 6 (Plasma cytokines & Summary) {.tabset}

### Fig 6a

```{r setSamples_corrVsSelf___Fig6a, ref.label=c('setSamples_olink')}
```


```{r AssociationWithTOA_prep___Fig6a}
 cor_method <- "spearman" ### "spearman", "pearson"
  corr_vsOther_vars <- c("TOA") ### specify which non-olink variables you would like to correlate with the olink variables
  corr_vsOther_pval <- 0.05 ## pvalue for annotation with stars & trimming of corrplot

  corr_group_var_names <- c("ALUM","ALFQA") ### these are are as in the spreadsheet, specify which values of the grouping variable you'd like to run corrplots for
  corr_timepoints <- c("Baseline","wk12_24h","wk13")
  
  ### suffix to name for output tab
  short_name_to_use <- "vsTOA"
  fig_name_to_use <- "Fig6a"
    assays_to_use  <- olink_assay_columns_new
    
    #### save values used to generate figure
    
   list_of_plotted[[paste0(fig_name_to_use)]] <- curr_phen_olink %>%
     ### remove basline
     dplyr::filter(Time_Point != "Baseline")
```


```{r AssociationWithTOA___Fig6a, ref.label=c('AssociationWithOther'), results='asis', fig.width=7, fig.height=6}
```



```{r Fig6a_prep}
cor_df_trim <- corVSTOA_df_prep  %>%
  ### trim to not baseline
  dplyr::filter(!grepl("Baseline",assay1)) %>%
  ### also make wider by time
  dplyr::mutate(timepoint = str_after_last(assay1,"__"),
                assay1 = str_before_last(assay1,"__")) %>%
  tidyr::pivot_wider(names_from = "timepoint", values_from = contains("_")) %>%
   ### trim to p < 0.05 at least one group
  dplyr::filter(if_any(starts_with("p_"), ~ . < corr_vsOther_pval))
 


##### save to list of stats & plotted
  list_of_plotted[[paste0(fig_name_to_use,"_sp")]] <- cor_df_trim
  list_of_stats[[paste0(fig_name_to_use)]] <- corVSTOA_df_prep  %>%
  ### trim to not baseline
  dplyr::filter(!grepl("Baseline",assay1)) %>%
  ### also make wider by time
  dplyr::mutate(timepoint = str_after_last(assay1,"__"),
                assay1 = str_before_last(assay1,"__")) %>%
   ### trim to p < 0.05 at least one group
  dplyr::filter(if_any(starts_with("p_"), ~ . < corr_vsOther_pval))
  


  cl_ratio <- 0.6
    pch_cex <- 2
    tl_cex <- 1.5

```

```{r vsTOAcorrplotting, results='asis', fig.width=7, fig.height=6}


          ##### separate out the R and p
          cor_mat <- cor_df_trim %>%
            column_to_rownames("assay1") %>%
            dplyr::select(contains("R_")) %>%
            ## remove _R from colnames
            `colnames<-`(c(str_after_last(colnames(.),"R_")))
          
                    ##### separate out the R and p
          cor_test_mat <- cor_df_trim %>%
            column_to_rownames("assay1") %>%
            dplyr::select(contains("p_")) %>%
            ## remove _R from colnames
            `colnames<-`(c(str_after_last(colnames(.),"p_")))
          
          
           
#### get row order
  vars_ann2 <- data.frame(cor_mat) %>%
    dplyr::arrange(desc(across(everything())))
  ordered_vars_to_use2 <- rownames(vars_ann2)
  
  
  #### get column order
  cols_ann <- data.frame("names"=colnames(cor_mat)) %>%
    dplyr::mutate(group = str_before_first(names, "_"),
                  tp = str_after_first(names,"_")) %>%
    dplyr::arrange(tp, desc(group))
  
  ordered_cols_to_use2 <- cols_ann$names


  ### need this in case have only since variable being compared
  cor_mat_for_plot2 <- as.matrix(cor_mat[ordered_vars_to_use2,ordered_cols_to_use2])
  rownames(cor_mat_for_plot2) <- ordered_vars_to_use2
  colnames(cor_mat_for_plot2) <- ordered_cols_to_use2

  cor_test_mat_for_plot2 <- as.matrix(cor_test_mat[ordered_vars_to_use2,ordered_cols_to_use2])
  rownames(cor_test_mat_for_plot2) <- ordered_vars_to_use2
  colnames(cor_test_mat_for_plot2) <- ordered_cols_to_use2


  if(length(which(cor_test_mat_for_plot2 <  as.numeric(corr_vsOther_pval))) == 0){


    corrplot::corrplot(cor_mat_for_plot2, method = 'square',
                       is.corr = FALSE, #since not symmetric
                       p.mat = cor_test_mat_for_plot2,
                       sig.level = as.numeric(corr_vsOther_pval), ## doesn't work if none are sig
                       #sig.level = 0.1,#set p < 0.05 to blank
                       insig='pch',
                       pch.cex = pch_cex, pch.col = 'black',tl.col = 'black',tl.cex=tl_cex,
                       ##legend
                       cl.ratio = cl_ratio, #cl.cex = 2,

                       #addCoef.col ='grey80',
                       #number.cex = 0.5, #add corr coeff to sig circles
                       #order = order, ## CANNOT ORDER ASSYMMETRIC
                       col = (COL2('PuOr',1000)), ##make orange NEGATIVE
                       col.lim = c(-1,1) ##set limits to be the same across plots
                       # title = vaccine_groups[i], mar=c(0,0,1,0) # http://stackoverflow.com/a/14754408/54964
    )


  }else{
    #### otherwise can use the stars
    corrplot::corrplot(cor_mat_for_plot2, method = 'square',
                       is.corr = FALSE, #since not symmetric
                       p.mat = cor_test_mat_for_plot2,
                       sig.level = as.numeric(corr_vsOther_pval),
                       insig='label_sig',  #set p < 0.05 to blank
                       pch.cex = pch_cex, pch.col = 'white',tl.col = 'black',tl.cex=tl_cex,
                       ##legend
                       cl.ratio = cl_ratio,#cl.cex = 2,
                       #addCoef.col ='grey80',
                       #number.cex = 0.5, #add corr coeff to sig circles
                       #order = order, ## CANNOT ORDER ASSYMMETRIC
                       col = (COL2('PuOr',1000)), ##make orange NEGATIVE
                       col.lim = c(-1,1) ##set limits to be the same across plots
                       # title = vaccine_groups[i], mar=c(0,0,1,0) # http://stackoverflow.com/a/14754408/54964
    )

  }

  cat("\n")
  cat(' \n \n') ### this is the key!
  
   
      
```



```{r scatterVsTOASetup___Fig6bd}
composite_fig_name <- "Fig6bd"
scatter_df <- data.frame("name"=c("Fig6b",
                                  "Fig6d"),
                         "y"=c("LTA__wk13",
                               "CCL8__wk12_24h"),
                         "trans_y"=c("identity",
                                     "identity"),
                         "dataset"=c("olink",
                                     "olink"))

```

```{r scatterplotsVsTOAPlot___Fig6bd, ref.label=c('scatterplotsVsTOAPlot'), results='asis', fig.height=4, fig.width=5}
```


```{r scatterVsOtherSetup___Fig6c}
composite_fig_name <- ""
scatter_df <- data.frame("name"=c("Fig6c"),
                         "x"=c("RectMucosa_DC10__Change"),
                         "trans_x"=c("identity"),
                         "y"=c("LTA__wk13"),
                         "trans_y"=c("identity"),
                         "dataset"=c("both"))

```

```{r scatterplotsVsOtherPlot___Fig6c, ref.label=c('scatterplotsVsOtherPlot'), results='asis', fig.height=4, fig.width=5}
```



### Fig 6e

Summary of variables different ALFQA vs Alum, associations with TOA





## Supp Figure 7 (Plasma cytokines & summary) {.tabset}

### Supp Fig7a

```{r SuppFig7a_prep}
fig_name_to_use <- "SuppFig7a"
cor_df_trim <- corVSTOA_df_prep  %>%
  ### also make wider by time
  dplyr::mutate(timepoint = str_after_last(assay1,"__"),
                assay1 = str_before_last(assay1,"__")) %>%
  tidyr::pivot_wider(names_from = "timepoint", values_from = contains("_")) 
### no further trimming
 


##### save to list of stats & plotted
list_of_plotted[[paste0(fig_name_to_use)]] <- curr_phen_olink
  list_of_plotted[[paste0(fig_name_to_use,"_sp")]] <- cor_df_trim
  list_of_stats[[paste0(fig_name_to_use)]] <- corVSTOA_df_prep  %>%
  ### also make wider by time
  dplyr::mutate(timepoint = str_after_last(assay1,"__"),
                assay1 = str_before_last(assay1,"__")) 
  


  cl_ratio <- 2
    pch_cex <- 1.2
    tl_cex <- 0.8

```

```{r vsTOAcorrplotting___SuppFig7a, ref.label=c('vsTOAcorrplotting'), results='asis', fig.width=7, fig.height=6}
```


### Supp Fig7b

```{r SuppFig7b_prep}
fig_name_to_use <- "SuppFig7b"
```

```{r setSamples_corrOlinkVsOther}
### generate curr_phen with BOTH olink and other assays
## first find animals that are in both datasets
overlap_animals <- intersect(unique(data_long$Sample_ID),unique(phen_olink$Sample_ID)) ### 41 animals



### curr_phen has rows = animals/time, columns = assays
curr_phen <- data_long %>%
  ## make simpler timepoint
  dplyr::mutate(Time_simp = ifelse(grepl("Wk",Time_Point),"WkPost",Time_Point),
                Time_Point = Time_simp,
                Time_simp = NULL,
                ### also make group with all caps
                group = toupper(Group)) %>%
  ## make wider
  unique() %>%
  ## remove duplicate rows with NA + number
  group_by(Sample_ID,Group, group,TOA,Time_Point,variable) %>%
  dplyr::reframe(across(everything(), ~ first(na.omit(.)))) %>%
  tidyr::pivot_wider(names_from = "variable",values_from = "value") %>%
  ### trim to overlapping animals
  dplyr::filter(Sample_ID %in% overlap_animals) %>%
  ####### add in olink
  dplyr::full_join(phen_olink %>%
                     ### trim to overlapping animals
                     dplyr::filter(Sample_ID %in% overlap_animals) %>%
                     ### combine variable & timepoint for olink assays
                     ## first need to make longer
                     tidyr::pivot_longer(cols = c(!!!rlang::syms(olink_assay_columns_new)),
                                         names_to = "variable",
                                         values_to = "value") %>%
                     dplyr::mutate(var_time = paste0(variable,"__",Time_Point)) %>%
                     ### remove anything with timepoint
                     dplyr::select(Sample_ID,var_time,value) %>%
                     tidyr::pivot_wider(names_from = "var_time",values_from = "value")
                   ,
                   by = c("Sample_ID"))

### save for later
curr_phen_both <- curr_phen
```



```{r AssociationOlinkVsOtherAssays_prep}
cor_method <- "spearman" ### "spearman", "pearson"
#### other vars will be the olink in this case
corr_vsOther_vars <- c(paste0(olink_assay_columns_new,"__Baseline"),
                       paste0(olink_assay_columns_new,"__wk12_24h"),
                       paste0(olink_assay_columns_new,"__wk13"))
corr_vsOther_pval_ann <- 0.05 ## pvalue for annotation with stars of corrplot
corr_vsOther_pval_cut <- 0.05 ## pvalue for trimming corrplot no quotes

corr_group_var_names <- c("ALUM","ALFQA") ### these are are as in the spreadsheet, specify which values of the grouping variable you'd like to run corrplots for
corr_timepoints <- c("Baseline","WkPost","Change")

### suffix to name for output tab
short_name_to_use <- "OtherVsOlink"
#### these are the NON-olink assays
assays_to_use_prep  <- colnames(curr_phen)[!grepl(paste(c("Sample_ID","Group","group","TOA","Time_Point"), collapse = '|'),colnames(curr_phen))]
assays_to_use  <- assays_to_use_prep[!grepl("__",assays_to_use_prep)]
```


```{r AssociationOlinkVsOther___SuppFig7b, ref.label=c('AssociationWithOther'), results='asis', fig.width=12, fig.height=6}
```

```{r SuppFig7b_printTables, results='asis'}

######### trim to only RM DC10

to_print_prep7b <- list_of_stats[[paste0(fig_name_to_use,"_",short_name_to_use)]] %>%
  dplyr::filter(assay1 %in% c("RectMucosa_DC10__WkPost","RectMucosa_DC10__Change")) %>%
  ### trim to only wk12_24h in olink
  dplyr::filter(grepl("wk12_24h",assay2)) %>%
  ### trim to only sig interactions
  dplyr::filter(p_ALFQA < 0.05) %>%
  
#### add MW dir for each assay
  ### first for RM
   dplyr::left_join(list_of_stats[[paste0("Fig3j_MW_2group")]] %>%
                      ## remove __ and make Wk generic
                      dplyr::mutate(rowname = ifelse(grepl("Wk",rowname),
                                                     paste0(str_before_last(rowname,"Wk"),"WkPost"),
                                                     rowname),
                                  #  rowname = str_replace_all(rowname,"__","_"),
                                    assay1_MW_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(rowname,assay1_MW_dir) %>%
                      ### rename rowname to assay1
                      dplyr::rename(assay1=rowname),
                    by = "assay1") %>%
  ### then for Olink
      dplyr::left_join(list_of_stats[[paste0("Fig5a_MW_2group")]] %>%
                      ## remove __ and make Wk generic
                      dplyr::mutate(#rowname = str_replace_all(rowname,"__","_"),
                                    assay2_MW_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(rowname,assay2_MW_dir) %>%
                      ### rename rowname to assay1
                      dplyr::rename(assay2=rowname),
                    by = "assay2") %>%
  
  
  #### add vsTOA for each assay
  
  ### first RM
      dplyr::left_join(list_of_stats[[paste0("Fig3j_vsTOA")]] %>%
                         ### trim to only this group
                         dplyr::select(assay1,!!!rlang::syms(colnames(.)[grepl(curr_group,colnames(.))])) %>%
                         ### make colnames generic
                         `colnames<-`(c(str_remove_all(colnames(.),curr_group))) %>%
                      ## remove __ 
                      dplyr::mutate(#assay1 = str_replace_all(assay1,"__","_"),
                                    assay1_vsTOA_dir = ifelse(p_ < 0.05,
                                                    ifelse(R_ > 0,
                                                           "pos",
                                                           "neg"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay1,assay1_vsTOA_dir),
                    by = "assay1") %>%
  
  ### then olink
      dplyr::left_join(list_of_stats[[paste0("Fig5a_vsTOA")]] %>%
                         ### trim to only this group
                         dplyr::select(assay1,!!!rlang::syms(colnames(.)[grepl(curr_group,colnames(.))])) %>%
                         ### make colnames generic
                         `colnames<-`(c(str_remove_all(colnames(.),curr_group))) %>%
                      ## remove __ 
                      dplyr::mutate(assay2 = assay1,
                                    assay2_vsTOA_dir = ifelse(p_ < 0.05,
                                                    ifelse(R_ > 0,
                                                           "pos",
                                                           "neg"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay2,assay2_vsTOA_dir),
                    by = "assay2") 

to_print <- to_print_prep7b %>%
 

     DT::datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           iDisplayLength = 50,
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
   print(htmltools::tagList(to_print))

        cat("\n")
      cat(' \n \n') ### this is the key!
      
      
      #### also add to plotted
      list_of_plotted[[paste0(fig_name_to_use)]] <- curr_phen_both ## full data used
      list_of_plotted[[paste0(fig_name_to_use,"_trimStats")]] <- to_print_prep7b
      list_of_stats[[paste0(fig_name_to_use)]] <- to_print_prep7b      
```


```{r scatterVsOtherSetup___SuppFig7cde}

composite_fig_name <- "SuppFig7cde"

scatter_df <- data.frame("name"=c("SuppFig7c",
                                  "SuppFig7d",
                                  "SuppFig7e"),
                         "x"=c("RectMucosa_DC10__Change",
                               "RectMucosa_DC10__Change",
                               "RectMucosa_DC10__Change"),
                         "trans_x"=c("identity",
                                     "identity",
                                     "identity"),
                         "y"=c("MMP12__wk12_24h",
                               "CCL13__wk12_24h",
                               "VEGFA__wk12_24h"),
                         "trans_y"=c("identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("both",
                                     "both",
                                     "both"))

```

```{r scatterplotsVsOtherPlot___SuppFig7cde, ref.label=c('scatterplotsVsOtherPlot'), results='asis', fig.height=4, fig.width=5}
```


### Supp Fig7f {.tabset}

```{r SuppFig7f_printTables_prep}
assays_to_keep <- "CCL8"
tp_to_keep <- "wk12_24h"
fig_name_to_use <- "SuppFig7f"
```

```{r olinkVsSelf_printTables, results='asis'}
to_print_prep_combo <- data.frame()
#### for each group
for(curr_group in c("ALUM","ALFQA")){
     cat("#### ",paste0(curr_group," "),"\n")
  cat('\n')
 

######### trim to only assays_to_keep vs other olink
#### start with olink vs self (includes inverted duplicates so only only have to trim on one assay)
to_print_prep7f <- list_of_stats[["Fig5a_vsSelf"]] %>%
  dplyr::filter(assay1 %in% assays_to_keep & tp == tp_to_keep) %>%
  ### remove self association
  dplyr::filter(assay1 != assay2) %>%
  ### trim to only sig interactions for at least one group
  dplyr::filter(!!rlang::sym(paste0(curr_group,"_p")) < 0.05) %>%
  
#### add MW dir for each assay
  ### first for assay1
   dplyr::left_join(list_of_stats[[paste0("Fig5a_MW_2group")]] %>%
                      ### trim to wk12_24h
                      dplyr::filter(grepl(tp_to_keep,rowname)) %>%
                      ### add dir
                      dplyr::mutate(assay1 = str_before_first(rowname,"__"),
                                    assay1_MWpost_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay1,assay1_MWpost_dir),
                    by = "assay1") %>%
    dplyr::left_join(list_of_stats[[paste0("Fig5a_MW_2group")]] %>%
                      ### trim to Baseline
                      dplyr::filter(grepl("Baseline",rowname)) %>%
                      ### add dir
                      dplyr::mutate(assay1 = str_before_first(rowname,"__"),
                                    assay1_MWBaseline_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay1,assay1_MWBaseline_dir),
                    by = "assay1") %>%
  ### then for assay2
      dplyr::left_join(list_of_stats[[paste0("Fig5a_MW_2group")]] %>%
                      ### trim to wk12_24h
                      dplyr::filter(grepl(tp_to_keep,rowname)) %>%
                      ### add dir
                      dplyr::mutate(assay2 = str_before_first(rowname,"__"),
                                    assay2_MWpost_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay2,assay2_MWpost_dir),
                    by = "assay2") %>%
      dplyr::left_join(list_of_stats[[paste0("Fig5a_MW_2group")]] %>%
                      ### trim to Baseline
                      dplyr::filter(grepl("Baseline",rowname)) %>%
                      ### add dir
                      dplyr::mutate(assay2 = str_before_first(rowname,"__"),
                                    assay2_MWBaseline_dir = ifelse(p_raw < 0.05,
                                                    ifelse(est > 0,
                                                           "up",
                                                           "down"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay2,assay2_MWBaseline_dir),
                    by = "assay2") %>%
  
  
  #### add vsTOA in this group
  
  ### first assay1 (assays_to_keep)
      dplyr::left_join(list_of_stats[[paste0("Fig5a_vsTOA")]] %>%
                         ### trim to only this group
                         dplyr::select(assay1,!!!rlang::syms(colnames(.)[grepl(curr_group,colnames(.))])) %>%
                         ### make colnames generic
                         `colnames<-`(c(str_remove_all(colnames(.),curr_group))) %>%
                         ### trim to only wk12_24h
                         dplyr::filter(grepl(tp_to_keep, assay1)) %>%
                      ## remove __ 
                      dplyr::mutate(assay1 = str_before_first(assay1,"__"),
                                    assay1_vsTOA_dir = ifelse(p_ < 0.05,
                                                    ifelse(R_ > 0,
                                                           "pos",
                                                           "neg"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay1,assay1_vsTOA_dir),
                    by = "assay1") %>%
  
  ### then olink
      dplyr::left_join(list_of_stats[[paste0("Fig5a_vsTOA")]] %>%
                         ### trim to only this group
                         dplyr::select(assay1,!!!rlang::syms(colnames(.)[grepl(curr_group,colnames(.))])) %>%
                         ### make colnames generic
                         `colnames<-`(c(str_remove_all(colnames(.),curr_group))) %>%
                          ### trim to only wk12_24h
                         dplyr::filter(grepl(tp_to_keep, assay1)) %>%
                      ## remove __ 
                      dplyr::mutate(assay2 = str_before_first(assay1,"__"),
                                    assay2_vsTOA_dir = ifelse(p_ < 0.05,
                                                    ifelse(R_ > 0,
                                                           "pos",
                                                           "neg"),
                                                    "not_sig")) %>%
                      ### trim to only dir & rowname
                      dplyr::select(assay2,assay2_vsTOA_dir),
                    by = "assay2") 

to_print <- to_print_prep7f %>%
 

     DT::datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           iDisplayLength = 50,
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
   print(htmltools::tagList(to_print))

        cat("\n")
      cat(' \n \n') ### this is the key!
      
      
      ### add to combo
      to_print_prep_combo <- rbind(to_print_prep_combo, to_print_prep7f %>%
                                     dplyr::mutate(Group = paste0(curr_group)) %>%
                                     dplyr::relocate(Group))
      
}
    cat("\n")
      cat(' \n \n') ### this is the key!
  
      
      
      
        #### also add to plotted & stats
      list_of_plotted[[paste0(fig_name_to_use)]] <-  list_of_plotted[[paste0("Fig5a")]] %>%
        ### trim to only this tp_to_use
        dplyr::select(group,Sample_ID,!!!rlang::syms(colnames(.)[grepl(tp_to_keep,colnames(.))]))
      list_of_plotted[[paste0(fig_name_to_use,"_trimStats")]] <- to_print_prep_combo
      list_of_stats[[paste0(fig_name_to_use)]] <- to_print_prep_combo      

      
```






## Supp Figure 8 (Plasma cytokines & summary) {.tabset}


```{r scatterVsTOASetup___SuppFig8abc}
composite_fig_name <- "SuppFig8abc"
scatter_df <- data.frame("name"=c("SuppFig8a",
                                  "SuppFig8b",
                                  "SuppFig8c"),
                         "y"=c("IL17F__wk13",
                               "CXCL8__wk13",
                               "EGF__wk13"),
                         "trans_y"=c("identity",
                                     "identity",
                                     "identity"),
                         "dataset"=c("olink",
                                     "olink",
                                     "olink"))

```

```{r scatterplotsVsTOAPlot___SuppFig8abc, ref.label=c('scatterplotsVsTOAPlot'), results='asis', fig.height=4, fig.width=5}
```




```{r scatterVsOtherSetup___SuppFig8d}
composite_fig_name <- ""
scatter_df <- data.frame("name"=c("SuppFig8d"),
                         "x"=c("CXCL8__wk13"),
                         "trans_x"=c("identity"),
                         "y"=c("EGF__wk13"),
                         "trans_y"=c("identity"),
                         "dataset"=c("olink"))

```

```{r scatterplotsVsOtherPlot___SuppFig8d, ref.label=c('scatterplotsVsOtherPlot'), results='asis', fig.height=4, fig.width=5}
```



### Supp Fig8e  {.tabset}

```{r SuppFig8e_printTables_prep}
assays_to_keep <- c("LTA","EGF","CXCL8","IL17F")
tp_to_keep <- "wk13"
fig_name_to_use <- "SuppFig8e"
```

```{r olinkVsSelf_printTables___SuppFig8e, ref.label=c('olinkVsSelf_printTables'), results='asis'}
```







## Data tables {.tabset}

### Raw Values 

```{r rawVals_print}
#### add with button
xfun::embed_file(file.path(input_dir,paste0("list_of_RAWdatasets.xlsx")), text = "Excel spreadsheet of RAW datasets")

```

### Plotted Values 

```{r plottedVals_print}

### sort tabs
list_of_plottedS <- list_of_plotted[order(names(list_of_plotted))]
##save
openxlsx::write.xlsx(list_of_plottedS,file.path(outputTABLES_dir,paste0("SourceDataFile.xlsx")))

#### add with button
xfun::embed_file(file.path(outputTABLES_dir,paste0("SourceDataFile.xlsx")), text = "Excel spreadsheet of plotted values")

```

### Stats {.tabset}

Mann-Whitney/Wilcoxon test between groups at each timepoint

Spearman correlation between TOA and assays, or among assays

```{r stats_print}
##save
openxlsx::write.xlsx(list_of_stats,file.path(outputTABLES_dir,paste0("list_of_stats.xlsx")))

#### add with button
xfun::embed_file(file.path(outputTABLES_dir,paste0("list_of_stats.xlsx")), text = "Excel spreadsheet of stats")

```



## Session info

```{r sessioninfo}
sessionInfo()
```
